time series exercises, week1

# Module 1: Tasks Basic Plots, Examples, and Fundamental Concepts

```{r}
rm(list = ls())
# install.packages("TSA")
library(magrittr)
library("TSA")
setwd("~/code/data-science/cheatsheet/source/includes/time-series/01-plots-fundamentals/tasks")
```

## 1. Plot stuff

- [ ] Load the Melbourne Airport annual rainfall data (MelbourneAnnualRainfall.csv) into R. 

```{r}
# ~old_todo~ change to data and data.ts for reusability
data <-
  read.csv("./raw-data/MelbourneAnnualRainfall.csv", header = FALSE)
rownames(data) <- seq(from = 1995, to = 2016)
class(data)
# Hint: “ts()” to convert the “data.frame” to timeSeries.
# ~old_todo~ doConvertTimeSeries()
data.ts <-
  ts(as.vector(data),
     start = 1995,
     end = 2016) # Convert to the TS object!
```

#### produce the time series plot for the annual rainfall series in Melbourne Airport.
```{r}
plot(data.ts, type = 'o', ylab = 'Amount of annual rainfall')
```

#### Do you think there is a trend in the series?
No obvious trend

## ~old_todo~ 2. Random values

#### - [x] Simulate a completely random process of `length 48` 
- [ ] with `independent, normal values`. 
- [ ] At this stage set the mean and variance of the normal distribution to `arbitrary values`.
```{r}
mu = 0
sigma = 10 #0.1 #sqrt{Var}
n = 48
data_random = rnorm(n, mu, sigma)
```

### - [ ] Display the time series plot without converting the generated data to a time series object 
```{r}
plot(data_random)
plot(ts(data_random))
plot.ts(data_random, 
  lty = 2 , 
  type = 'o', 
  ylab = 'Normal distributed process') 
#without converting the generated data to a time series object
```

### ~old_todo~ check #converting the generated data to a time series object
```{r}
plot(ts(data_random) , type = 'o', ylab = 'Normal distributed process') #converting the generated data to a time series object
```

- [ ] Does it look “random?”

~old_todo~ wk1q2. Plot with random values
- [ ] Display the time series plot without converting the generated data to a time series object 
  - [ ] and by converting the generated data to a time series object.
  - [ ] Does it look “random?”
- [ ] Repeat this exercise several times with a new simulated dataset at each time. Does your view on the randomness of series change from one simulation to the other?
- [ ] Repeat this exercise with standard deviations of 0.1 and 10. What is the effect of variance of series on the resulting time series plot?


## ~old_todo~ 3. Chi Squared exercise
3. Simulate a completely random process of `length 48` with independent, `chi-square distributed values`, `each with 2 degrees of freedom`.
- Display the time series plot. Does it look “random?”
- Repeat this exercise several times with a new simulated dataset at each time. Does your view on the randomness of series change from one simulation to the other?
- Compare your results with those obtained in the second task.

```{r}
nu = 2
n = 48
data_chisq = rchisq(n, nu) # generate data
data_chisq.ts <- ts(data_chisq) #convert to a time series object
plot(
  data_chisq.ts, 
  type = 'o',
  ylab = 'Chi-squre distributed process')   
```

4. Labelling monthly
- Load the Melbourne Airport **monthly** rainfall data (MelbourneMonthlyRainfall.csv into R.

```{r}
data <-
  MelbourneMonthlyRainfall <- 
    read.csv(
    "raw-data/MelbourneMonthlyRainfall.csv",
    header = FALSE)
# MelbourneMonthlyRainfall
```

- [ ] Construct a time series plot in a blank graphics window. Is the seasonal effect is clearly seen from the time series plot?


- [ ] Construct a time series plot in a blank graphics window with monthly plotting symbols. 

```{r}
# ~old_todo~ WRONG: learn this. This breaks it, because data is wrong
rownames(data) <- seq(from = 1995, to = 2016)
colnames(data) <-
  c("jan",
    "feb",
    "mar",
    "apr",
    "may",
    "jun",
    "jul",
    "aug",
    "sep",
    "oct",
    "nov",
    "dec")
class(data)

data <-
  ts(
    as.vector(as.matrix(data)),
    start = c(1995, 1),
    end = c(2016, 12),
    frequency = 12
  )
data
# Now the data read into R in an incorrect order!!! Check with the original CSV file!
```

### Right: This breaks it, because data is wrong
```{r}
# Right: This breaks it, because data is wrong

# Read the raw data again
data <- MelbourneMonthlyRainfall

# Take the transpose of the matrix with t() to put in the correct order!!!
data_fix_month_order.ts <-
  ts(
    as.vector(as.matrix(t(
      data
    ))),
    start = c(1995, 1),
    end = c(2016, 12),
    frequency = 12
  )
data_fix_month_order.ts %>% head()

class(data_fix_month_order.ts)
```

### Seasonal plot
```{r}
# --- open a new graphics window in Mac OS ---
# x11()

# --- open a new graphics window in Windows ---
# win.graph(width = 4.875, height = 2.5, pointsize = 8)

plot(data_fix_month_order.ts, type = 'o', ylab = 'Amount of monthly rainfall')
points(
  y = data_fix_month_order.ts,
  x = time(data_fix_month_order.ts),
  pch = as.vector(season(data_fix_month_order.ts))
)
```

```{r}
season_data = season(data_fix_month_order.ts)
season_data %>% head(12)
```

- Is there any improvement in the ease of identification of the seasonal effect when the symbols are super-imposed onto the plot?

Sure
