---
title: "wk1 Fundamental Concepts"
output: html_notebook
---

# Module 2: Tasks Basic operations and Decompositions

```{r}
working_directory <- '/Users/phil/code/data-science-next/datasets/forecasting'
setwd(working_directory)

library(packagr)
packages <- c(
  # 'dylnm',
  'dLagM', 'forecast', 'expsmooth', 'TSA', 'Hmisc', 'car', 'AER', 'readr', 'tseries', 'lubridate')
packagr(packages) # alpha package to install & load packages

source('/Users/phil/code/data-science-next/uni/common/utils-forecasting.R')
source('/Users/phil/code/data-science-next/uni/common/utils.R')

```

The aims of the tasks given here are to:

- [ ] make you familiar with basic operations in time series analysis and their implementation in R, and
- [ ] foster your practical ability to conduct time series decompositions and inferring the results arising from time series decompositions.

## Exercises:

### 1. Load the monthly commercial landings dataset of US-NMFS from `Metric_Tons` column of the file `NMFS_Landings.csv` into the workspace of R. Do the following tasks over the landings series;


```{r}
data_filename <- 'NMFS_Landings.csv'
nmfs <- NMFS_Landings <- read_csv(
  data_filename,
  col_names = TRUE, 
)
```

```{r}
# Re-order df by year and month

nmfs$Date <- paste(nmfs$Year, nmfs$Month) %>% parse_date_time('ym')
nmfs <- arrange(nmfs, Date)

class(NMFS_Landings)
# nmfs %>% head() # 1991 Jan
# nmfs %>% tail() # 2015 Dec
```

```{r}
nmfs.ts <- ts(
  nmfs$Metric_Tons, 
  start = c(1991, 1),
  end = c(2015, 12),
  frequency = 12
)
nmfs.ts %>% head()

```

- [ ] Display and infer time series, ACF and PACF plots.
```{r}
plot(nmfs.ts)

acf(nmfs.ts)

pacf(nmfs.ts)
```
Possible seasonality

- [ ] Apply a suitable transformation to this series. 
```{r}
confidenceInterval(nmfs.ts)
```
lambda includes 0

```{r}
nmfs.ts.log = log(nmfs.ts)
doDiffAndPlot(
  nmfs.ts.log, 0, T, T,
  fig_nums("log_transform", "Log transform"))
```


- [ ] Does it work for overcoming the problem with this series?

No



- [ ] Take the ordinary and/or seasonal difference to overcome the general problem with this series. Does it work?

```{r}
# if we have to log transform it first
nmfs.ts.log.decompose <- decompose(nmfs.ts.log, type="additive")
plot(nmfs.ts.log.decompose)
```

```{r}
# if not
nmfs.ts.decompose <- decompose(nmfs.ts, type="additive")
plot(nmfs.ts.decompose)
```


- [ ] Decompose this series into trend, seasonal and remainder components and infer the results.

```{r}

```

- [ ] Plot the naïve forecasts for this series.

```{r}

```



### 2. The file [global_warming.csv](https://climate.nasa.gov/vital-signs/global-temperature/ ) 

- includes annual mean and 5 year mean series between 1880 and 2016 
- showing the change in global surface temperature relative to 1951-1980 average temperatures. 
- See the visualisation in the source page mentioned above. 

Load the annual mean series into the workspace of R and do the following tasks:

```{r}
data_filename <- 'global_warming.csv'
global_warming = read_csv(
  "/Users/phil/code/data-science-next/datasets/forecasting/global_warming.csv", 
  col_names = TRUE, 
)
class(global_warming)
# global_warming %>% head()
```

- [ ] Display and infer time series, ACF and PACF plots.

```{r}

```

- [ ] Apply a suitable transformation to this series. Does it work for overcoming the problem with this series?

```{r}

```

- [ ] Take the ordinary and/or seasonal difference to overcome the general problem with this series. Does it work?

```{r}

```

### 3. Load the consumer price inflation in Australia

- [inflation.csv](https://www.rba.gov.au/statistics/tables/xls/g01hist.xls?v=2017-07-09-21-49-05 ) into the workspace of R. 

- [ ] Display and infer time series, ACF and PACF plots.

```{r}
data_filename <- 'inflation.csv'
inflation = read_csv(
  "/Users/phil/code/data-science-next/datasets/forecasting/inflation.csv", 
  col_names = TRUE, 
)
# class(inflation)
# inflation %>% head()
```

- [ ] Apply a suitable transformation to this series. Does it work for overcoming the problem with this series?

```{r}

```

- [ ] Take the ordinary and/or seasonal difference to overcome the general problem with this series. Does it work?

```{r}

```

- [ ] Decompose this series into trend, seasonal and remainder components and infer the results.

```{r}

```

- [ ] Plot the naïve forecasts for this series.

```{r}

```

