---
title: "R Notebook"
output: html_notebook
---

```{r}

setwd("~/code/tldr/data-science/stats-math1324/03week/")

library(psych)
library(mosaic)

library(readr)

Mobilephone <- read_csv("Mobilephone - Mobilephone.csv")
head(Mobilephone_Mobilephone)

mobile_phone$usr <- factor(mobile_phone$usr, 
                    levels=c("R",
                             "S",
                             "U"), 
                    labels=c("Rural",
                    	       "Suburban",
                    	       "Urban"))

mobile_phone$SmartPhone <-factor(mobile_phone$SmartPhone, levels=c(1,2,3),
                                	labels=c("Cell, smartphone", 
                                	         "Cell, not a smartphone",
                                	         "No cell"))

```

Use the cross tabulation to answer the following questions. Write out each question as a probability statement, e.g. Pr(A).

What’s the probability a person was from a rural area?

```{r}

tally (~SmartPhone | usr , data="mobile_phone", margin= TRUE  )

```

What’s the probability that a person’s residency area was missing?

```{r}

```

What’s the probability a person had a smartphone and lived in an urban area?

```{r}

```

What’s the probability a person owned a mobile phone?

```{r}

```

What’s the probability a person had a smartphone given they were from a rural area?

```{r}

```

What’s the same probability, but for a person who lived in a suburban area?

```{r}

```

What should be concluded about smartphone ownership and area of residence?

```{r}

```





