---
title: "Week 11 - Class Worksheet"
subtitle: "Special Operations: Dealing with date/time and character variables"
author: "Dr. Anil Dolgun"
date: "16/05/2018"
output: 
  html_document:
    number_sections: false
    theme: cerulean
    highligt: pygments
---

Class activities
```{r}
library("purrr")

weekCount <- (10)

weeklymeetings <- function(weekCount) {
 (today() + dweeks(1:10)) %>%   return()
}

weeklymeetings()

  map(.x = weekCount, .f = weeklymeetings())



weeklymeetings(10)

Map(f = ,)
today() + dweeks(1)


````



## Required Packages 

The following packages and the function will be required or may come in handy.

```{r}
library(readr)
library(dplyr)
library(lubridate)
library(stringr)

```

## Exercises 


### Pollution Data

The following exercises 1-4 will be based on [Pollution.csv](../data/Pollution.csv) data from Kaggle <https://www.kaggle.com/nicapotato/pollution-in-atchison-village-richmond-ca/data>. Variables are self explanatory however it is expected to do checks on the type of the data and use the suitable transformations if necessary. 
 
Here is a quick look of the `Pollution` data:


1. Check the structure of the `Date` variable. Convert it to date format using the appropriate function. After the conversion it should look like this:

```{r}
pollution <- read.csv("~/code/tldr/data-science/datasets/Pollution.csv", header=FALSE)
head(pollution)


```


2. Create year, month, day, hour and minute columns with using `mutate()`. Create a new column with combining year, month and day. 

```{r}

pollution %>% mutate(time = make_datetime()

pollution_mutate <- pollution %>% mutate(departure = make_datetime(year, month, day, hour, minute)) 


```


3. Create a new column with Today's date and name it `newtime`. Add 2 years on the `Date` variable. Check the difference between `Date` and `newtime` variable in weeks, round it to 2. Check the duration of the `year` variable using appropriate `duration` function.

`difftime` function

```{r}

lubridate

date() + years()

lubr

my_age_days <- difftime(time1 = date(), time2 ="17082013")


```


4. Create a sequence of time starting from 6:00 am 1st of May 2015 ending 7:00 am 1st of October 2015 by hour. Check the length of this sequence. Create a subset of pollution data with the same length of the sequence and bind them. 


```{r}

```


### NYC Jobs Data

The following exercises 5-10 will be based on NYC Jobs data set [NYC_jobs.csv](../data/NYC_jobs.csv) from <https://data.cityofnewyork.us>. You can find the data description in <https://data.cityofnewyork.us/City-Government/NYC-Jobs/kpav-sd4t>.  Variables are self explanatory however it is expected to do checks on the type of the data and use the suitable transformations if necessary. Here is a quick look of the few variables of NYC Jobs data:



5. Check the structures of the variables `Posting Date`, `Posting Updated` and `Process Date`. If any of the variables are in character format, convert it to date format. Hint: To do so you will need to use chartr() first for `Posting Date` to replace T with a blank.

```{r}

```



6. Paste `Job Id` and `Title Code No` with a seperator "-", name this column as new_id. Use both `paste()` and `str_c()`.

```{r}

```


7. Convert characters to lower for `Agency`, `Business title` and `Civil Service Title`. If you are looking for a challenge use `mutate_at()` function.

```{r}

```


8. Trim white spaces for both ends and find string length for `Job Description` variable.

```{r}

```


9. Pad the `Job Id` variable with a leading `1` and pick 7 for the width.

```{r}

```


10. Extract the first characater of the `Salary Frequency`.

```{r}

```


## Finished?

If you have finished the above tasks, work through the weekly list of tasks posted on the Canvas announcement page.

<center>
**[Return to Data Preprocessing Website](../index.html)**
</center>
