---
title: "MATH1324 Introduction to Statistics Assignment 2"
author: "Phil Steinke s3725547"
output: html_notebook
---

# Modelling Body Measurements

### 1. Problem Statement  [Plain text]:
_Write a clear and concise problem statement that guides your investigation. Explain which variable you choose and outline the approaches taken for normal distribution fitting._

TODO:

- This paper examins the wrist girth of male and females from the dataset [Heinz G, Peterson LJ, Johnson RW, Kerk CJ. 2003. Exploring Relationships in Body Dimensions. Journal of Statistics Education 11(2).] (https://ww2.amstat.org/publications/jse/v11n2/datasets.heinz.html). 

- Differences in body dimensions between male and females against the normal disntribution

- It examines the factors `wri.di` wrist girth (in cm) and `sex` (male or female)

###  2. Load Packages:
```{r}
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(base))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library("mosaic", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library"))
setwd("~/code/tldr/data-science/stats-math1324/assignment2")
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("latticeExtra"))

```

### 3. Data: Import the body measurements data and prepare it for analysis.

```{r}
body_measurements <- read_excel("data/bdims.csv (1).xlsx",
    sheet = "bdims.csv") %>%
  as.data.frame()
```

#### Tidy it up

_define the variable sex as a factor and define labels for it) using R functions._

```{r}
wrist_measurements <-
  select(
    body_measurements,
    wri.gi,
    sex) %>%
  as.data.frame()

colnames(wrist_measurements) <-
  c("wrist.girth", "sex")

wrist_measurements <-
  mutate(
    wrist_measurements,
    sex=factor(
      sex,
      labels=c("female", "male"))) %>%
  print()

```

## 4. Summary Statistics: Calculate descriptive statistics

_Calculate descriptive statistics (i.e., mean, median, standard deviation, first and third quartile, interquartile range, minimum and maximum values) of the selected measurement grouped by sex._

## Summary of wrist measurements for all sexes
```{r}
wrist_summary <- wrist_measurements$wrist.girth

summary(wrist_summary)

summary_mean <-
  mean(wrist_summary) %>%
  print()

summary_sd <-
  sd(wrist_summary) %>%
  print()

IQR(wrist_summary) # Interquartile Range

```

## Male Summary Statistics
```{r}
male <- subset(
  wrist_measurements,
  sex == "male"    
  )$wrist.girth

summary(male)

male_mean <-
  mean(male) %>%
  print()

male_sd <-
  sd(male) %>%
  print()

IQR(male) # Interquartile Range

```

## Female Summary Statistics
```{r}
female <- subset(
  wrist_measurements,
  sex == "female"  
  )$wrist.girth

summary(female)

female_mean <-
  mean(female) %>%
  print()

female_sd <-
  sd(female) %>%
  print()

IQR(female) # Interquartile Range

```



## 5. Distribution Fitting: histogram with normal distribution overlay

- Compare the empirical distribution of selected body measurement to a normal distribution separately in men and in women

```{r}

histogram(
  x = wrist_summary,
  type="density",
  main = "Wrist girth (cm) of both sexes and its normal distribution",
  col=rgb(0,0,0,alpha=0.0),
  xlim=c(12,21),
  ylim = c(0,0.50),
  breaks = 40,
  xlab = "wrist girth (cm)",
  ylab = "bar"
  )

plotDist(
  "norm",  
  mean = summary_mean, 
  sd = summary_sd, 
  kind= "density",
  type = "b",
  col = "red",
  lwd=2,
  add=TRUE,
  xlab = "wrist girth (cm)",
  ylab = "bar"
  ) %>%
  return()

```

- Then you will use R to summarise the empirical distribution of body measurement separately in men and women and compare it to a normal distribution. You need to do this visually by plotting the histogram with normal distribution overlay.

## summarise the empirical distribution of body measurement


## compare it to a normal distribution
- by plotting the histogram with normal distribution overlay


### for men
```{r}

histogram(
  x = male,
  type="density",
  main = "male wrist girth (cm) against normal distribution",
  col=rgb(0,0,0,alpha=0.0), xlim=c(12,21), ylim = c(0,0.80),
  breaks = 40,
  xlab = "wrist girth (cm)",
  ylab = "Density Histogram\n with Normal Fit"
  )

plotDist(
  "norm",
  mean = summary_mean,
  sd = summary_sd,
  kind= "density",
  type = "b",
  col = "red",
  lwd=2,
  add=TRUE,
  xlab = "wrist girth (cm)",
  )

plotDist(
  "norm",  
  mean = male_mean,
  sd = male_sd,
  kind= "density",
  type = "b",
  col = "blue",
  lwd=2,
  add=TRUE,
  xlab = "wrist girth (cm)",
  ) %>%
  return()

```

### for females
```{r}

histogram(
  x = female,
  type="density",
  main = "female wrist girth (cm) compared to normal distribution",
  col=rgb(0,0,0,alpha=0.0), xlim=c(12,21), ylim = c(0,0.80),
  breaks = 40
  )

plotDist(
  "norm",  
  mean = summary_mean,
  sd = summary_sd,
  kind= "density",
  type = "b",
  col = "red",
  lwd=2,
  add=TRUE) %>%
  return()

plotDist(
  "norm",  
  mean = female_mean,
  sd = female_sd,
  kind= "density",
  type = "b",
  col = "pink",
  lwd=2,
  add=TRUE) %>%
  return()

```

## Interpretation:


### What insight has been gained from the investigation?

- The mean of males is `~17.2cm` and females are `~15cm`
- Median of male is `~17.1cm` and female is `15cm`
- The girth of male wrists are larger than female wrists by an average of `2cm`

### Discuss the extent to how your theoretical normal distribution fits the empirical data

- The data fits a similar distribution 
- There are very few female cases above `17cm`
- Male, female and both sexes fit a normal bell curve



### Make recommendations regarding the modelling of this body measurement

- Further analysis to check against other distributions
