---
title: "5/9/2018MATH1324 Introduction to Statistics Week 10 - Class"
output: html_notebook
---


```{r}
drunks <- read.csv("~/code/tldr/data-science/datasets/stats-math1324-csv/Drunkswalk - Drunkswalk.csv")
head(drunks)
library("mosaic")

```

MATH1324 - Class Worksheet
Week 10 - Chi-square Goodness of Fit Test and Chi-square Test of Association
Complete the following exercises after watching the Module 8 video demostration.

Datasets: Drunkswalk.csv (see Data Repository)
Gender and Drunken Stumbling
Activity Description
Reanalyse the Drunkswalk.csv data, but this time 

## determine whether gender (0 = females , 1 = males) was associated with the probability of stumbling to the side of a drunk’s dominant hand.

```{r}

drunks$Gender <- factor(drunks$Gender,levels = c(0,1), labels=c("hello ladies", "drunk gents"))

drunks$Match <- factor(drunks$Match,levels = c(0,1), labels=c("Non-dominant", "Dominant"))
tally(~Match, data = drunks, margins = TRUE)

tally(~Match | Gender, data = drunks, margins = TRUE, format="proportion")

tally(~Match | Gender, data = drunks, margins = TRUE, format="proportion")

```


Produce a crosstabulation of the counts and proportions of males and females stumbling to their dominant and nondominant side.

Visualise the association between gender and whether a drunk stumbled to their dominant or nondominant side using a clustered bar chart. Comment on the association.

Perform an appropriate hypothesis test to determine if a drunk’s direction of stumble, to either the side of their dominant or nondominant hand, was associated with gender. Use the 0.05 level of significance.
State the Null and Alternate hypothesis for the appropriate hypothesis test. State any assumptions.

Report the test statistic, df, and p-value from the results of the hypothesis test. Comment on any assumption checked.

Use the p-value to make a decision about the Null hypothesis.

Conclude whether or not the results of the hypothesis test were statistically significant and draw a conclusion in the context of the example.
Fair Die
Activity Description
The following table shows the outcomes of 1200 rolls of a die. Are the die fair?

Outcome
1
2
3
4
5
6
Count
207
189
224
193
215
172
 

Load the count data into RStudio as a numeric vector named count.

Create another numeric vector named fair which contains the expected proportion for each outcome of a die assuming the die are fair.

Perform an appropriate hypothesis test to determine if die are fair. Use the 0.05 level of significance.
State the Null and Alternate hypothesis for the appropriate hypothesis test. State any assumptions.

Report the test statistic, df, and p-value from the results of the hypothesis test. Comment on any assumption checked.

Use the p-value to make a decision about the Null hypothesis.

Conclude whether or not the results of the hypothesis test were statistically significant and draw a conclusion in the context of the example.
R Hints
Use this example R code if you get stuck:
Gender and Drunken Stumbling
> Drunkswalk$Match<-factor(Drunkswalk$Match,
levels = c(0,1),labels=c("Non-dominant",
"Dominant"))

> Drunkswalk$Gender<-factor(Drunkswalk$Gender, levels = c(0,1), 
labels = c("Female","Male"))

> tally(~Match | Gender, data = Drunkswalk,margins = TRUE)

> tally(~Match | Gender, data = Drunkswalk,
margins = TRUE,format="proportion")

> table2<-tally(~Match | Gender, data = Drunkswalk,format="proportion")

> barplot(table2, 
        main = "Stumble Side by Gender",ylab="Proportion within Gender",
        ylim=c(0,1),legend=rownames(table2),beside=TRUE,
        args.legend=c(x = "topright",horiz=TRUE,title="Stumble Direction"),
        xlab="Gender")
> grid()

> xchisq.test(tally(~Match | Gender, data = Drunkswalk))

Fair Die
> count<-c(207,189,224,193,215,172)

> fair<-rep(1/6,6)

> xchisq.test(count,p = fair)


 non-dominant sideusing a clustered bar chart.Exercise 5
5/9/2018MATH1324 Introduction to Statistics Week 10- Class Worksheetfile://rmit.internal/USRHome/el7/E24447/Downloads/Week_10_Class_Worksheet_Answers%20(1).html5/6a- State the Null and Alternate hypothesis for the appropriate hypothesis test. State any assumptions.# H0: There is no association between gender and whether a drunk stumbles to their dominant hand side# HA: There is an association between gender and whether a drunk stumbles to their dominant hand side# Assumption: No more than 25% of expected cell counts are below 5b- Report the  statistic, , and -value from the results of the hypothesis test. Comment on any assumption checked.Perform an appropriate hypothesis test to determine if a drunk’s direction of stumble, to either the side of theirdominant or non-dominant hand, was associated with gender. Use the 0.05 level of significance.Exercise 6χ2dfp
5/9/2018MATH1324 Introduction to Statistics Week 10- Class Worksheetfile://rmit.internal/USRHome/el7/E24447/Downloads/Week_10_Class_Worksheet_Answers%20(1).html6/6chi2 <-chisq.test(tab1)chi2## ##  Pearson's Chi-squared test with Yates' continuity correction## ## data:  tab1## X-squared = 4.7654e-31, df = 1, p-value = 1chi2$expected##               ##                  Female     Male##   Non-dominant 24.18543 19.81457##   Dominant     58.81457 48.18543# There are no cells with expected counts below 5.c- Use the -value to make a decision about the Null hypothesis.# p = 1, fail to reject H0.d- Conclude whether or not the results of the hypothesis test were statistically significant and draw a conclusion in the context of the example.# The Chi-square test of association was not statistically significant. There was no evidence of an association between gender and whether a drunk stumbled to the side of their dominant hand. Males and females stumbled to the side of their dominant hand a similar proportion of the time.Return to Introduction to Statistics Website (https://astral-theory-157510.appspot.com/secured/index.html)p
