---
title: "R Notebook"
output: html_notebook
---

```{r}
hmm, I keep getting this random problem with saving as RData

```{R}

# Random data.frame
foo_dataframe <- data.frame(runif(10), runif(10), runif(10))
save(foo_dataframe, file = "data/foo.RData")
bar <- load(file = "data/foo.RData")
identical(foo_dataframe, bar) # [1] FALSE

foo_matrix <- matrix(runif(10))
attributes(foo_matrix)
save(foo_matrix, file = "data/foo_matrix.RData")
bar_matrix <- load(file = "data/foo_matrix.RData")
attributes(bar_matrix)
identical(foo_matrix, bar_matrix) # [1] FALSE
help(load)
# Got it working with readRDS
saveRDS(foo_dataframe, file = "data/foo.rds")
bar <- readRDS("data/foo.rds")
identical(foo_dataframe, bar) # TRUE

unlink(".RData")

```


```{r}

# p <- plot_ly(family_violence, x = , y =  type = "bar")

p <- plot_ly (
      x = family_violence$`Age Range`,
      y = "family_violence[3]",
      type = "bar",
      mode = "markers" )
p

# api_create(p, filename = "name-of-my-plotly-file") save-online
# Cheatsheet https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf
# reference https://plot.ly/r/reference/

```
