---
title: "wk1 Fundamental Concepts"
output: html_notebook
---

wavelet( layers = )

# Module 1: Tasks

```{r}
working_directory <- '/Users/phil/code/data-science-next/datasets/forecasting'
setwd(working_directory)

library(packagr)
packages <- c(
  # 'dylnm',
  'dLagM', 'forecast', 'expsmooth', 'TSA', 'Hmisc', 'car', 'AER',
  'readr', 'tseries')
packagr(packages) # alpha package to check, install and load packages

source('/Users/phil/code/data-science-next/uni/time-series/common/utils.r')
```


## Fundamental Concepts

The aims of the tasks given here are to:

- make you familiar with data type conversions required in this course,
- foster your practical ability to produce main time series displays,
- strengthen your ability to draw inferences from the plots and hypothesis tests.
- Working individually or in small groups, please complete the following exercises.

## 1. Load the number of earthquakes per year magnitude 7.0 or greater between 1900 and 1998 in US given in earthquakes.csvPreview the document file (Source: http://datamarket.com/) into R and do the following tasks

```{r}
data_filename <- 'earthquakes.csv'
earthquakes = read_csv(
  "/Users/phil/code/data-science-next/datasets/forecasting/earthquakes.csv", 
  col_names = TRUE, 
)
# class(earthquakes)
# earthquakes %>% head()
```

### Convert the given data into the `ts`

```{r}
earthquakes.ts = ts(earthquakes$Number, start = 1900) 
earthquakes.ts %>% head()
# earthquakes.ts %>% class()
```

### Display and infer time series and ACF plots.

```{r}
# doDiffAndPlot(earthquakes, 0, TRUE, FALSE)
plot(
  earthquakes.ts,
  type = 'l'
)

acf(earthquakes.ts)

pacf(earthquakes.ts)
```

### What do you think stationarity of this series? Please answer by conducting a unit root test.

- Not stationary 

```{r}
k1 = ar(earthquakes.ts)$order
print(k1)

adf.test(earthquakes.ts, k = k1)
```

```{r}
k2 = trunc(12*((length(earthquakes.ts)/100)^(1/4)))
print(k2)

adf.test(earthquakes.ts, k = k2)
```

```{r}
k3 = trunc(4*((length(earthquakes.ts)/100)^(1/4)))
print(k3)

adf.test(earthquakes.ts, k = k3)
```



```{r}
PP.test(earthquakes.ts, lshort = TRUE)
```

```{r}
PP.test(earthquakes.ts, lshort = FALSE)
```

```{r}
adfTest(earthquakes.ts)
```

```{r}
eacf(earthquakes.ts)
```

## 2. Load mean monthly flow recorded for Piper’s Hole River between 1953 and 1981 given in `flow.csv`. Preview the document file (Source: http://datamarket.com/) into R and do the following tasks

```{r}
data_filename <- 'flow.csv'
flow = read_csv(
  "/Users/phil/code/data-science-next/datasets/forecasting/flow.csv", 
  col_names = TRUE, 
)
flow %>% head()
flow %>% tail()

flow$Month %>% length()
```

### Convert the given data into the `ts`

```{r}
flow.ts = ts(
  flow$Flow,
  start = c(1953, 1),
  end = c(1981, 12),
  frequency = 12
  ) 

data <-
  ts(
    as.vector(as.matrix(data)),
    start = c(1995, 1),
    end = c(2016, 12),
    frequency = 12
  )

  flow.ts
```

### Display and infer time series and ACF plots.

```{r}
plot(flow.ts)
```

```{r}
acf(flow.ts)
```

### What do you think stationarity of this series? Please answer by conducting a unit root test.

```{r}
k1 = ar(flow.ts)$order
print(k1)
adf.test(flow.ts, k = k1)
```

```{r}
k2 = trunc(12*((length(flow.ts)/100)^(1/4)))
print(k2)
adf.test(flow.ts, k = k2)
```

```{r}
test  = adf.test(flow.ts)
test$parameter

```

```{r}
PP.test(bonds, lshort = TRUE)
PP.test(bonds, lshort = FALSE)


```


## 3. Load mean monthly median aerosol optical depth at Cape Grim between 1986 and 2014 given in aerosol.csv Preview the documentfile (Source: https://data.gov.au/) into R and do the following tasks

### Convert the given data into the ts
### Display and infer time series and ACF plots.
### What do you think stationarity of this series? Please answer by conducting a unit root test.

## 4. Simulate a random walk process of length 100.

### Display and infer time series and ACF plots.
### What do you think stationarity of this series? Please answer by conducting a unit root test.