---
title: "MATH2349 Semester 1, 2018"
author: "Phil Steinke s3725547"
subtitle: Assignment 1
output:
  html_document:
    df_print: paged
  html_notebook: default
---

## Setup

Install and load the necessary packages to reproduce the report here:

```{r, echo = TRUE, warnings = FALSE}

# This is a chunk where you can load the necessary packages required to reproduce the report 

# Here are some example packages, you may add others if you require  

library(readr) # Useful for importing data
library(foreign) # Useful for importing SPSS, SAS, STATA etc. data files
library(rvest) # Useful for scraping HTML data
library(knitr) # Useful for creating nice tables

setwd("~/Documents/code/data-science-apps/data-preprocessing/assignment1")


```

## Data Description

Locate an open source of data from the web. This can be a tabular, spreadsheet data (i.e., .txt, .csv, .xls, .xlsx files), data sets from other statistical software (i.e., SPSS, SAS, Stata etc.), or you can scrape HTML table data.

Some sources for open data are provided below, but I encourage you to find others:

* [http://www.abs.gov.au/](http://www.abs.gov.au/)

* [https://www.data.vic.gov.au/](https://www.data.vic.gov.au/) 	

* [http://www.bom.gov.au/](http://www.bom.gov.au/)

* [https://www.kaggle.com](https://www.kaggle.com) 

As a minimum, your data set should include:

* one numeric variable.
* one qualitative (categorical) variable.

There is no limit on the number of observations and number of variables. But keep in mind that when you have a very large data set, it will increase your reading time. **A clear description of data and its source should be provided in this section.** 


## Read/Import Data

- Read/Import the data into R, 
- then save it as a data frame. 

You can use Base R functions or readr, xlsx, readxl, foreign, rvest packages for this purpose. 

You must also provide the R codes with outputs (i.e. `head()` of the data set) that you used to import/read/scrape the data set. You can provide the R codes with outputs using R chunks like this:

```{r}

# This is an R chunk for importing the data. Provide your R codes here:

# Read/Import the data into R
ok_cupid_data <- readRDS("data/parsed_data_public.rds")

# save it as a dataframe
# TODO:

# You must also provide the R codes with outputs (i.e. `head()` of the data set) that you used to import/read/scrape the data set
head(ok_cupid_data)



ok_cupid_data_head <- readRDS("data/parsed_data_public.rds", function(x) head(x,2))
head(ok_cupid_data_head)

```


Explain everything that you do in this step using regular text outside the R chunks. You may use bulleted lists like this:


Here is an example of bulleted list:

* Bullet 1
* Bullet 2
* ....



## Inspect and Understand


Inspect the data frame and variables using R functions. As a minimum, you should:

* check the dimensions of the data frame.
* check the data types (i.e., character, numeric, integer, factor, and logical) of the variables in the data set.
* check the levels of factor variables, rename/rearrange them if required.
* check the column names in the data frame, rename them if required.


Provide your R codes with outputs and explain everything that you do in this step.


```{r}

# This is a chunk where you inspect the types of variables, data structures, check the attributes in the data.



```

## Subsetting I

Subset the data frame using first 10 observations (include all variables). Then convert it to a matrix. Provide the R codes with outputs and explain everything that you do in this step. 



```{r}

# This is a chunk to subset your data and convert it to a matrix 


```

If you get an error, you won't be able to preview the notebook. To avoid this, you may use `eval = FALSE` option inside the R chunk like this:

```{r, eval=FALSE}

# With eval = FALSE option knitr won't run the R code but will display your codes on the notebook. 


```

Do not forget to explain why you got this error.

## Subsetting II


Subset the data frame including only first and the last variable in the data set, save it as an R object file (.RData). Provide the R codes with outputs and explain everything that you do in this step. 

```{r}

# This is a chunk to subset your data and convert it to an R object file 


```


## IMPORTANT NOTE: 

This report must be uploaded to Turnitin as a PDF with your code chunks showing. The easiest way to achieve this is to Preview your notebook in HTML (by clicking Preview) → Open in Browser (Chrome) → Right click on the report in Chrome → Click Print and Select the Destination Option to Save as PDF. 


**Feel free to DELETE the instructional text provided in the template.** If you have any questions regarding the assignment instructions and the R template, please post it on [Slack](https://math2349.slack.com) under the `#Assignment1` channel. 

<br>
<br>
