---
title: "R Notebook"
output: html_notebook
---

```r
install.packages("readr")
install.packages("rvest")
install.packages("xlsx")

setwd("~/Documents/code/data-science-apps/data-preprocessing/import-scraping-export-week2")

dir.create("foo2")

getwd()

setwd("./.")

getwd()

iris <- read_csv("iris.csv")
View(iris)

str(iris)

View(iris1)

head(iris)

# 3. Import csv version of Population dataset (population.csv) using Base R functions.

base_population <- read.csv("population.csv")
head(base_population)

# 4.  Repeat exercise 3, this time use readr functions.

install.packages("readr")
library(readr)
readr_population <- read_csv("population.csv")
head(readr_population)

# 5. Import spss version of Population data (population.sav) using foreign package.

install.packages("foreign")
library(foreign)
```



# 3. Import csv version of Population dataset (population.csv) using Base R functions.

```{r}

base_population <- read.csv("population.csv")
head(base_population)
```

# 4.  Repeat exercise 3, this time use readr functions.


```{r}
install.packages("readr")
library(readr)
readr_population <- read_csv("population.csv")
head(readr_population)
```

# 5. Import spss version of Population data (population.sav) using foreign package.

```{r}
install.packages("foreign")
library(foreign)

population_save <- read.spss("population.sav", to.data.frame = TRUE)

ls()

head(population_save, 10)

```

# Population - Migration Data
Exercise 6 is based on population-migration dataset (population-migration.xls) which contains two sheets named “Population Density and Regional” and “Inter-regional Migration” in .xls format. “Inter-regional Migration” sheet contains 1564 rows of timeseries data between 2009 to 2016. This data set is also taken from OECD located at http://stats.oecd.org/Index.aspx?DataSetCode=REGION_DEMOGR. The variables for population-migration data (“Inter-regional Migration” sheet) contains:

Region [Character]: Region name

Territory Level and Typology [Factor]: Description of the region (Country, Large regions, Small regions,… )

2009, 2010, ...,2016 [Numeric]: Migration (All persons inflows minus outflows)

**Import population-migration dataset (population-migration.xls) using xlsx or readxl functions. (Hint: Use  sheetName argument.)**

```r

install.packages("readxl")
library(readxl)

xl <- "population-migration.xls""

population_migration_density <- read_excel(xl, 
    sheet = "Population Density and Regional")

View(population_migration_density)

```

## From the lecture:
```r
install.packages(readxl)
library(readxl)
# read in xlsx worksheet using a sheet index or name

iris13<- read_excel("iris.xlsx", sheet = "iris")
# read in xlsx worksheet and change variable names by skipping the first row
# and using col_names to set the new names

iris14<- read_excel("iris.xlsx", sheet = "iris", skip = 1, col_names = paste ("Var", 1:6))
```

## From the docs:
```r
# Usage
read_excel(path, sheet = NULL, range = NULL, col_names = TRUE,
col_types = NULL, na = "", trim_ws = TRUE, skip = 0, n_max = Inf,
guess_max = min(1000, n_max))

```

## Most Popular Baby Names by Sex and Mother’s Ethnic Group, New York City Data
The following exercises (exercise 7-9) is based on Popular Baby Names, NYC dataset which contains 22,035 rows of data recorded in 2011. This data set is taken from NYC Open Data which is located at https://data.cityofnewyork.us/Health/Most-Popular-Baby-Names-by-Sex-and-Mother-s-Ethnic/25th-nujf containing the following variables:

BIRTHYEAR [Integer]: Year of birth

GENDER [Character]: Gender

ETHNICITY [Character]: Mother’s ethnicity (categories: HISPANIC, ASIAN AND PACIFIC ISLANDER, WHITE NON HISPANIC, BLACK NON HISPANIC)

NAME [Character]: Child first name

COUNT [Integer]: Count

RANK [Integer]: Rank
The url for the csv file is located at https://data.cityofnewyork.us/api/views/25th-nujf/rows.csv?accessType=DOWNLOAD

## 7 Use the url for Popular Baby Names, NYC dataset to import in R.

```r

setwd("~/Documents/code/data-science-apps/data-preprocessing/import-scraping-export-week2")

popular_baby_names_url <- "https://data.cityofnewyork.us/api/views/25th-nujf/rows.csv?accessType=DOWNLOAD"

popular_baby_names <- read.csv(popular_baby_names_url, stringsAsFactors = FALSE)

head(popular_baby_names)

```

## 8 After you import the dataset save it as a .csv file.

```r

setwd("~/Documents/code/data-science-apps/data-preprocessing/import-scraping-export-week2")

popular_baby_names_url <- "https://data.cityofnewyork.us/api/views/25th-nujf/rows.csv?accessType=DOWNLOAD"

popular_baby_names <- read.csv(popular_baby_names_url, stringsAsFactors = FALSE)

head(popular_baby_names)

```


## 9 Repeat exercise 8, this time save it as .Rdata.

```r

```





