---
title: "Week 4 - Class Worksheet"
output: html_notebook
---


Tidy and Manipulate: Part I - Tidy
Dr. Anil Dolgun
Last updated: 19 March, 2018
Required Packages
The following packages will be required or may come in handy.

```{r}
library(tidyr)
library(dplyr)
library(readr)
library(readxl)
library(knitr)
```

## Exercises
# Family Incidents data set
The following exercises (exercise 1-4) are based on family incidents (family_incidents.csv)[http://rare-phoenix-161610.appspot.com/secured/data/family_incidents.csv] data set. This data set is taken from Crime Statistics Agency https://www.crimestatistics.vic.gov.au/family-violence-data-portal/download-data-tables containing 84 observation from Local government areas about family incident rate per 100,000 population by police region from 2012 to 2017. The variables for family_incidents.csv data set consists of:
Police Region [Character]: Police regions (levels: North West Metro, Eastern, Southern Metro, Western and where Total2 is for unknown geographical location).

```{r}
setwd("data")

family_incidents <- read_csv("family_incidents.csv")
View(family_incidents)

```

Local Government Area [Character]: 79 Local government areas of Victoria there are totals per region.

2012-2013, ..., 2016-2017: The year that data was collected.
First six observations of the family_incidents data set:
Police region	Local government area	2012-13	2013-14	2014-15	2015-16	2016-17
North West Metro	Banyule	940.9	974.1	1060.8	1030.6	1014.3
North West Metro	Brimbank	986.8	1084.3	1240.5	1343.3	1105.0
North West Metro	Darebin	1020.8	1131.3	1063.6	1071.9	1052.4
North West Metro	Hobsons Bay	899.5	1065.7	1137.0	1205.7	1075.9
North West Metro	Hume	1390.7	1540.3	1554.9	1538.0	1478.6
North West Metro	Maribyrnong	956.8	1042.4	997.6	1089.4	891.8
Read in the family_incidents.csv data set using a suitable function. Check out the classes of each variable and convert Police Region column into factor, use ordered argument then check its’ levels.

Choose the suitable action for family_incident data set.

 a) We need to combine multiple columns into a single column using unite() function

 b) We need to gather those columns into a new pair of variables using gather() function - This one

 c) We need to split the variables since multiple variables are stored in one column using separate() function

 d) We need to transform data from long format to wide format using spread() function

# 2. Tidy the family_incidents data set into the form below.

```{r}
# From the lecture
table4a %>%
gather(`1999`, `2000`, key = "year", value = "cases")

family_incidents_



```

Police region	Local government area	year	cases
North West Metro	Banyule	2012-13	940.9
North West Metro	Brimbank	2012-13	986.8
North West Metro	Darebin	2012-13	1020.8
North West Metro	Hobsons Bay	2012-13	899.5
Using family_incidents data set, change the separator of the year column from "-" to "/" using  separate() and unite() functions.
Primary Health Network data set
The following exercises (exercise 5-8) are based on the PHN data set (PHN.xls). This data set has the percentage of children who are immunised against Polio and Hepatitis B in Australia by age group and location in October 2017 - September 2017. PHN data set is taken from http://www.immunise.health.gov.au/internet/immunise/publishing.nsf/Content/current-phn-immunisation-coverage-data with 93 observations, containing variables:
PHN Number [Character]: ID of the Primary Health Network Area

PHN Name [Character]: Name of the Primary Health Network Area

Age Group [Character]: Factor with Levels 12-<15 Months, 24-<27 Months, 60-<63 Months

%Polio [Numeric]: Percentage of children immunised against Polio

%HEP [Numeric]: Percentage of children immunised against Hepatitis B
First six observations of the PHN data set:
PHN Number	PHN Name	Age Group	%Polio	%HEP
PHN101	Central and Eastern Sydney	12-<15 Months	94.20931	94.09876
PHN102	Northern Sydney	12-<15 Months	94.75332	94.58254
PHN103	Western Sydney	12-<15 Months	94.07182	93.99051
PHN104	Nepean Blue Mountains	12-<15 Months	94.94118	94.94118
PHN105	South Western Sydney	12-<15 Months	93.85654	93.92611
PHN106	South Eastern NSW	12-<15 Months	95.44135	95.54297
Read in the PHN.xls data set using a suitable function. Check out the classes of each variable and convert Age Group column into factor, use ordered argument then check its levels.

Assume that we want to calculate the mean percentage as a column of all age groups of children immunised against Polio and Hepatitis B. Before calculating the mean, select the tidy functions that we need in order to form the data set as:

PHN Number	PHN Name	Vaccination Type	12-<15 Months	24-<27 Months	60-<63 Months
PHN101	Central and Eastern Sydney	%HEP	94.09876	95.37775	0.00000
PHN101	Central and Eastern Sydney	%Polio	94.20931	95.54502	92.50768
PHN102	Northern Sydney	%HEP	94.58254	95.11925	0.00000
PHN102	Northern Sydney	%Polio	94.75332	95.56295	92.16040
PHN103	Western Sydney	%HEP	93.99051	95.89433	0.00000
PHN103	Western Sydney	%Polio	94.07182	95.99414	94.36424
 a) unite(), gather()

 b) split(), unite()

 c) gather(), spread()

 d) separate(), split()

Tidy the PHN data set into the form in exercise 6. Once you tidy, use rowMeans function from baseR package to calculate the mean percentage of children immunised and save this average as a vector. (Hint: Remember how to select columns as a sequence i.e., df[,3:5].) Repeat this calculation by combining your tidy code and calculating the mean in one line whilst saving it as a vector.

Bonus exercise: Use round function from base R package to round the values to 2 decimals places in PHN data set.

Finished?
If you have finished the above tasks, work through the weekly list of tasks posted on the Canvas announcement page.

Return to Data Preprocessing Website