
```{r include=FALSE}
# Packages:

# devtools::install_git('https://gitlab.com/botbotdotdotcom/packagr')
library(packagr)
packages <- c(
  'dLagM', 'forecast', 'expsmooth', 'TSA', 'Hmisc', 'car', 'AER',
  'readr', 'tseries', 'lubridate', 'stringr', 'testthis', 'captioner', 'urca')
packagr(packages) # alpha package to check, install and load packages

source('/Users/phil/code/data-science-next/uni/common/utils-forecasting.R')

# Config Options:
# data %>% head(1) # Initial date: Jan-04 (from original date column)
file_name = '~/code/data-science-next/datasets/forecasting/ASX_data.csv'
tsStart = c(2004, 1)
# data %>% tail(1) # Final date: May-17 (from original date column)
tsEnd = c(2017, 5)
default_ylab = 'ASX data'
default_xlab = 'Year'
featuresData <- list('asx.ts', 'gold.ts', 'oil.ts', 'copper.ts')

# Load data:
data <- read_csv(
  file_name,
  col_names = TRUE,)
names(data) <- c('asx', 'gold', 'oil', 'copper')
```


##### Create timeSeries objects

```{r}

with12Frequency.ts <- ts(
  data$oil,
  start = tsStart,
  frequency = 12
); 

with6Frequency.ts <- ts(
  data$oil,
  start = tsStart,
  frequency = 6
); 

with1Frequency.ts <- ts(
  data$oil,
  start = tsStart,
  frequency = 1
); 


noFrequency.ts <- ts(
  data$oil,
  start = tsStart
); 

decompose12Frequency.ts <- decompose(with12Frequency.ts, type="multiplicative") 
plot(decompose12Frequency.ts$seasonal)

decompose6Frequency.ts <- decompose(with6Frequency.ts, type="multiplicative") 
plot(decompose6Frequency.ts$seasonal)

# decompose1Frequency.ts <- decompose(with1Frequency.ts, type="multiplicative") 
# plot(decompose1Frequency.ts$seasonal)

plot(noFrequency.ts)

isSeasonal(with12Frequency.ts)
isSeasonal(with6Frequency.ts)
isSeasonal(with1Frequency.ts)
isSeasonal(noFrequency.ts)



# decomposeNoFrequency.ts <- decompose(noFrequency.ts, type="multiplicative") 
# plot(noFrequency.ts)
