
```{r}
library(dLagM)
library(forecast)

working_directory <- '/Users/phil/code/data-science-next/datasets/forecasting'
data = read.csv("global_warming_CO2.csv")

class(data)
head(data)

warming = ts(data$Warming,start = 1959)
co2 = ts(data$CO2,start = 1959)
data.ts = ts(data[,2:3],start = 1959)
```

#### We can plot them individually

```{r}
plot(warming, ylab = "Relative temperature change", xlab = "Year", main = "Global warming series")
plot(co2, ylab = "Parts per million (ppm) ", xlab = "Year", main = "Mid-tropospheric carbon dioxide series")
```

#### We can plot them in the same frame but the scales will be problemtatic and we won't be able to 

```{r}
see warming series cearly.
plot(data.ts, plot.type="s",col = c("blue", "red"), main = "Global warming and CO2 series")
legend("topleft",lty=1, text.width = 11, col=c("blue","red"), c("Y series", "X series"))
```
#### We can scale and center both series to see in the same plot clearly
```{r}
data.scaled = scale(data.ts)
plot(data.scaled, plot.type="s",col = c("blue", "red"), main = "Global warming and CO2 series")
legend("topleft",lty=1, text.width = 11, col=c("blue","red"), c("Y series", "X series"))
```
#### The correlation between warming and CO2 series is very high
```{r}
cor(data.ts)
```
#### Fit a finite DLM with a lag length of 4 and do the diagnostic checking.
#### No need to supply ts object. The function looks for vectors.
```{r}
model1 = dlm( x = as.vector(co2) , y = as.vector(warming), q = 4 )
summary(model1)
checkresiduals(model1$model)
```

#### Find the finite lag length based on AIC and BIC for this data
#### We can change  q and compare AIC/BIC values
```{r}
for ( i in 1:10){
  model1.1 = dlm( x = as.vector(co2) , y = as.vector(warming), q = i )
  cat("q = ", i, "AIC = ", AIC(model1.1$model), "BIC = ", BIC(model1.1$model),"\n")
}
```
#### We go for the smallest AIC/BIC.
#### Also we can use finiteDLMauto function as follows
```{r}
finiteDLMauto( x = as.vector(co2) , y = as.vector(warming), q.min = 1, q.max = 10, 
              model.type = "dlm", error.type ="AIC", trace = TRUE)

finiteDLMauto( x = as.vector(co2) , y = as.vector(warming), q.min = 1, q.max = 10, 
               model.type = "dlm", error.type ="BIC", trace = TRUE)
```
#### In the outpt "q - k" shows the lag length.
```{r}
model1.2 = dlm( x = as.vector(co2) , y = as.vector(warming), q = 1 )
summary(model1.1)
checkresiduals(model1.1$model)
```
#### Fit a polynomial DLM of order 2. Display the estimates of original parameters and their 
significance tests.
```{r}
model2 = polyDlm(x = as.vector(co2) , y = as.vector(warming) , q = 2 , k = 2 , show.beta = TRUE)
summary(model2)
checkresiduals(model2$model)
```
#### Find forecasts for the CO2 levels 406.31, 407, and 415. Here 406.31 ppm is the latest measurement in May 2017. 
#### Plot the forecasts along with the original warming series.
```{r}
model1.2Frc = dLagM::forecast(model1.2, x = c(406.31, 407, 415), h = 3)
model2Frc = dLagM::forecast(model2, x = c(406.31, 407, 415), h = 3)

plot(ts(c(as.vector(warming), model1.2Frc$forecasts), start = 1959),type="o",col="Blue", ylim= c(-0.3, 1.7), 
     ylab = "Warming", xlab = "Year", main="Global warming series with three years ahead forecasts")      
lines(ts(c(as.vector(warming), model2Frc$forecasts), start = 1959),col="red",type="o")
lines(ts(as.vector(warming), start = 1959),col="black",type="o")
legend("topleft",lty=1, pch = 1, text.width = 11, col=c( "blue", "red","black"), c("Dlm", "Polynomial", "Warming"))
```

#### Fit the Koyck model. Find and plot the forecasts for the case given above.

```{r}
model3 = koyckDlm(x = as.vector(co2) , y = as.vector(warming))
summary(model3)
checkresiduals(model3$model)
```

#### Fit autoregressive distributed lag models and choose one of the models.
```{r}
model4 = ardlDlm(x = as.vector(co2) , y = as.vector(warming), p = 1 , q = 1 )
summary(model4)
checkresiduals(model4$model)

for (i in 1:5){
  for(j in 1:5){
    model4.1 = ardlDlm(x = as.vector(co2) , y = as.vector(warming), p = i , q = j )
    cat("p = ", i, "q = ", j, "AIC = ", AIC(model4.1$model), "BIC = ", BIC(model4.1$model),"\n")
  }
}
```
p =  1 q =  1 gives the smallet AIc and BIC
Find forecasts for the CO2 levels 406.31, 407, and 415. Here 406.31 ppm is the latest 
measurement in May 2017. 

#### Plot the forecasts along with the original warming series.
```{r}
model3Frc = dLagM::forecast(model3, x = c(406.31, 407, 415), h = 3)
model4Frc = dLagM::forecast(model4, x = c(406.31, 407, 415), h = 3)

plot(ts(c(as.vector(warming), model1.2Frc$forecasts), start = 1959),type="o",col="Blue", ylim= c(-0.3, 1.7), 
     ylab = "Warming", xlab = "Year", main="Global warming series with three years ahead forecasts")      
lines(ts(c(as.vector(warming), model2Frc$forecasts), start = 1959),col="red",type="o")
lines(ts(c(as.vector(warming), model3Frc$forecasts), start = 1959),col="green",type="o")
lines(ts(c(as.vector(warming), model4Frc$forecasts), start = 1959),col="purple",type="o")
lines(ts(as.vector(warming), start = 1959),col="black",type="o")
legend("topleft",lty=1, pch = 1, text.width = 11, col=c( "blue", "red", "green", "purple", "black"), c("Dlm", "Polynomial", "Koyck", "ARDL", "Warming"))
```