# Module 5: Tasks
# Exponential smoothing methods

```{r}
library(TSA)
library(car)
library(dynlm)
library(Hmisc)
library(forecast)
library(xts)
```

# Task 1

In this task we will deepen the analysis of monthly retail unleaded fuel prices series for Melbourne, Vic. recorded between May 1998 and June 2017 (See http://www.aaa.asn.au/latest-fuel-prices/). Please load the data series from the file `fuelPriceData.csv` into the workspace of R and do the following tasks:


```{r}
price <- read.csv(  '~/code/data-science/datasets/forecasting/seaLevel.csv',)
price = ts(price[,2], start=c(1998,4),frequency = 12)
```

- [ ] Display the series using suitable descriptive plots and draw inferences about the mechanism behind the series.
```{r}
plot(price,ylab='Fuel price',xlab='Year',main = "Time series plot of the monthly retail unleaded fuel prices series.")
points(y=price,x=time(price), pch=as.vector(season(price)))
```

- [ ] Fit a set of suitable classical exponential smoothing methods and choose one of the models as the best fitting one within the set of tentative models considering diagnostic checking as well.

```{r}
fit1.ses <- ses(price, initial="simple", h=5) 
summary(fit1.ses)
checkresiduals(fit1.ses)
```

```{r}
fit2.holt <- holt(price, initial="simple", h=5) # Let the software estimate both alpha and beta
summary(fit2.holt)
checkresiduals(fit2.holt)

fit3.holt <- holt(price, initial="simple", exponential=TRUE, h=5) # Fit with exponential trend
summary(fit3.holt)
checkresiduals(fit3.holt)

fit4.holt <- holt(price, damped=TRUE, initial="simple", h=5) # Fit with additive damped trend
summary(fit4.holt)
checkresiduals(fit4.holt)
```

```{r}
plot(fit1.ses, type="l", ylab="Fuel price", xlab="Year", 
     fcol="white", plot.conf=FALSE)
lines(fitted(fit1.ses), col="blue") 
lines(fitted(fit2.holt), col="red")
lines(fitted(fit3.holt), col="green")
lines(fitted(fit4.holt), col="cyan")
lines(fit1.ses$mean, col="blue", type="l") 
lines(fit2.holt$mean, col="red", type="l")
lines(fit3.holt$mean, col="green", type="l")
lines(fit4.holt$mean, col="brown", type="l")
legend("topleft", lty=1, col=c("black","blue","red","green","cyan"), 
       c("Data","SES", "Holt's linear trend", "Exponential trend","Additive damped trend"))

```

```{r}
fit5.hw <- hw(price,seasonal="additive", h=5*frequency(price))
summary(fit5.hw) 
checkresiduals(fit5.hw)

fit6.hw <- hw(price,seasonal="additive",damped = TRUE, h=5*frequency(price))
summary(fit6.hw)
checkresiduals(fit6.hw)

fit7.hw <- hw(price,seasonal="multiplicative", h=5*frequency(price))
summary(fit7.hw)
checkresiduals(fit7.hw)

fit8.hw <- hw(price,seasonal="multiplicative",exponential = TRUE, h=5*frequency(price))
summary(fit8.hw)
checkresiduals(fit8.hw)
```

### State Space

- [ ] Apply state-space versions of exponential smoothing methods focused on the previous task and select the best fitting model using model selection criteria and diagnostic checking results.

```{r}

fit1.etsA = ets(price, model="ANN")
summary(fit1.etsA)
checkresiduals(fit1.etsA)

fit1.etsM = ets(price, model="MNN")
summary(fit1.etsM)
checkresiduals(fit1.etsM) # Least autocorrelated - BG test

fit2.etsA = ets(price, model="AAN")
summary(fit2.etsA)
checkresiduals(fit2.etsA)

fit2.etsM = ets(price, model="MAN", damped = TRUE)
summary(fit2.etsM)
checkresiduals(fit2.etsM)

fit3.etsA = ets(price, model="AAA")
summary(fit3.etsA)
checkresiduals(fit3.etsA)

fit3.etsM = ets(price, model="MAA")
summary(fit3.etsM)
checkresiduals(fit3.etsM) # By visual checks residuals of this model is less correlated.

fit4.etsM = ets(price, model="MAM")
summary(fit4.etsM)
checkresiduals(fit4.etsM)

fit5 = ets(price)
summary(fit5)
checkresiduals(fit5) # Autofit actually finds the fit1.etsM which is the least autocorrelated - BG test
```

### TODO: Model Fitting

### Forecast
- [ ] Decide which method performs better in terms of model fitting and forecasts?
```{r}
plot(forecast(fit1.etsM), ylab="Fuel price",plot.conf=FALSE, type="l", xlab="Year")
```

```{r}
plot(forecast(fit3.etsM), ylab="Fuel price",plot.conf=FALSE, type="l", xlab="Year")
```

# Task 2
High-quality measurements of global sea level have been made since late 1992 by satellite altimeters. The series in file `SeaLevel.csv` shows the global mean sea level (mm) between January 1993 and August 2016 (http://www.cmar.csiro.au/sealevel/sl_hist_last_decades.html).

```{r}
seaLevel <- read.csv("~/Documents/MATH1307_Forecasting/Tasks/Task 5/seaLevel.csv")
seaLevel = ts(seaLevel$SeaLevel, start=c(1993,1), frequency = 12)
```

- [ ] Display the series using suitable descriptive plots and draw inferences about the mechanism behind the series.

```{r}
plot(seaLevel, ylab="The global mean sea level (mm)", xlab= "Time", main="Time series plot for global mean sea level series.")
# Seasonality, trend and changing variance
```

- [ ] Fit a set of suitable classical exponential smoothing methods and choose one of the models as the best fitting one within the set of tentative models considering diagnostic checking as well.

```{r}
seaLevel2 = log(seaLevel+abs(min(seaLevel))+0.1)
plot(seaLevel2, ylab="The global mean sea level (mm)", xlab= "Time", main="Time series plot for global mean sea level series.")
# Log transformation did not help.
```

```{r}
fit1.sea = hw(seaLevel,seasonal="additive", h=5*frequency(seaLevel))
summary(fit1.sea) 
checkresiduals(fit1.sea)

fit2.sea = hw(seaLevel,seasonal="additive", damped = TRUE, h=5*frequency(seaLevel))
summary(fit2.sea) # Best fit, best time series plot for residuals
checkresiduals(fit2.sea)

# We add a constant to be able to fit multiplicatie model with negative or zero values
fit3.sea = hw(min(seaLevel+50),seasonal="multiplicative", h=5*frequency(seaLevel))
summary(fit3.sea) 
checkresiduals(fit3.sea)

fit4.sea = hw((seaLevel+50),seasonal="multiplicative",damped = TRUE, h=5*frequency(seaLevel))
summary(fit4.sea) 
checkresiduals(fit4.sea)

fit5.sea = hw((seaLevel+50),seasonal="multiplicative",damped = FALSE, exponential = TRUE, h=5*frequency(seaLevel))
summary(fit5.sea) 
checkresiduals(fit5.sea)
```

### State Space
- [ ] Apply state-space versions of exponential smoothing methods focused on the previous task and select the best fitting model using model selection criteria and diagnostic checking results.

```{r}
fit1.sea.ets = ets(seaLevel, model = "AAA")
summary(fit1.sea.ets)
checkresiduals(fit1.sea.ets)

fit2.sea.ets = ets(log(seaLevel+50), model = "AAA", damped = TRUE)
summary(fit2.sea.ets) # MASE = 0.4553113
checkresiduals(fit2.sea.ets)

fit3.sea.ets = ets((seaLevel+50), model = "MAA")
summary(fit3.sea.ets)
checkresiduals(fit3.sea.ets)

fit4.sea.ets = ets((seaLevel+50), model = "MAA", damped = TRUE)
summary(fit4.sea.ets)
checkresiduals(fit4.sea.ets)

fit5.sea.ets = ets((seaLevel+50), model = "MMM")
summary(fit5.sea.ets)
checkresiduals(fit5.sea.ets)

fit6.sea.ets = ets((seaLevel+50), model = "MMM", damped = TRUE)
summary(fit6.sea.ets)
checkresiduals(fit6.sea.ets)
```

- [ ] Use the automatic model fitting feature of the forecast package and compare the results with those found in the previous two bullets in terms of the selected model, diagnostic check results and the accuracy of forecasts and model fitting.

```{r}
fit.auto = ets((seaLevel+50))
summary(fit.auto) # BEST MASE = 0.4613409 but better time series plot of residuals, non of the models is able to remove autocorrelations in residuals.
```

```{r}
checkresiduals(fit.auto)
```

```{r}
plot(forecast(fit.auto), ylab="Sea level",plot.conf=FALSE, type="l", xlab="Year")
```
