---
title: "MATH1318 Time Series"
author: "Phil Steinke s3725547"
subtitle: Assignment 2 - Semester 1, 2019
output:
  html_document:
    highlight: kate
    theme: paper
  pdf_document:
    fig_caption: yes
---

### Data

Egg depositions (in millions) of age-3 [Lake Huron Bloaters (Coregonus hoyi)](https://en.wikipedia.org/wiki/Coregonus_hoyi) between years 1981 and 1996 are available in BloaterLH dataset of FSAdata package.

### Goals

- [x] Analyse the egg depositions of Lake Huron Bloasters 
- [x] Use the analysis methods covered in the **modules 1 – 7** of MATH1318 Time Series Analysis
- [x] Choose the best model among a set of possible models for this dataset
- [x] Give forecasts of egg depositions for the next 5 years

#### R code

```{r Setup, message=FALSE, warning=FALSE, warnings=0}
cat("\014") # clear everything
setwd("./.")
dataFilename <- "raw-data/eggs.csv"
startYear = 1981 # get start year
endYear = 1996 # get end year

devtools::install_git('https://gitlab.com/botbotdotdotcom/packagr')
library(packagr)

packages <- c('FitAR','fUnitRoots','lmtest','readr','tidyr','TSA','TTR','MASS','purrr','captioner')
packagr(packages) # alpha package to check, install and load packages
data <- read_csv(dataFilename)
# convert to timeseries, ignoring year column:
data.ts <- ts(as.vector(data[,2]), start = startYear, end = endYear) 
# source('~/code/data-science/uni/time-series/common/sort.score.R')
# sort.score function from Haydar
sort.score <- function(x, score = c("bic", "aic")){
  if (score == "aic"){
    x[with(x, order(AIC)),]
  } else if (score == "bic") {
    x[with(x, order(BIC)),]
  } else {
    warning('score = "x" only accepts valid arguments ("aic","bic")')
  }
}
data.ts.raw <- data.ts # make a copy for forecasting
```
```{r data-check}
# class(data) -> ts
# data %>% dim() # 16 x 2 
data.ts %>% dim() # 16 x 1 because we removed the year column
```

```{r diff-functions}
default_ylab = 'Lake Huron Bloaters'
default_xlab = 'Year'
fig_nums <- captioner()

doDiffAndPlot <- function(df.ts, diffCount, showPlot = T, showEacf=F) {
  ifelse(
    diffCount != 0, (df.ts = diff(df.ts, differences = diffCount)),
    'there is no diff\n')
  paste('diff: ', diffCount) %>% print()
  order = ar(diff(df.ts))$order
  paste('order count (k): ', order)
  diffAdfTest = adfTest(df.ts, lags = order, title = NULL, description = NULL)
  p <- diffAdfTest@test$p.value
  paste(
    'adf p-value:',
    p, (ifelse(p < 0.05, '< 0.05 significant', '> 0.05 insignificant')
    )) %>% print()
  if(showPlot) {
    par(mfrow = c(1,3))
    plot(
      df.ts, 
      type='o', 
      xlab=default_xlab, 
      # ylab=fig_nums(figureName, 
      #   paste('diff #', diffCount, ' of ', default_ylab, sep = '')
      # )
    )
    acf(df.ts)
    pacf(df.ts)
    par(mfrow=c(1,1))
  }
}
```

```{r initial-data}
# The initial ts data, without any transformation:
fig_nums('initial ts data', 'initial timeseries data with no transformation') %>% cat()
doDiffAndPlot(data.ts, 0, T, F) # p = 0.4455 lag = 1 - has a trend = nope
```

- Plot has visual positive trend
- Small dataset, so it is hard to visually observe any significant seasonality
- ACF and PACF have a waveform (similar to sin/cosine wave)
- Has a clear lag of 1 in ACF and PACF

## Confidence interval

```{r confidence-interval, warning=FALSE}
# check the confidence interval of BoxCox
fig_nums("confidence-interval", paste('Confidence Interval for ', default_ylab, sep = '')) %>% cat()
boxcoxCi <- BoxCox.ar(data.ts, method = "yule-walker")$ci
```

- The confidence interval `0.1 - 0.8` does not include 0, so we _cannot_ do a log transform

```{r test-normality}
doTestNormality <- function(df) {
  qqnorm(df)
  qqline(df, col = 2)
  shapiro.test(df)
}
lambda <- (max(boxcoxCi)-min(boxcoxCi))/2 
# lambda=0.35 ~midpoint betwen confidence interval (0.8-0.1)/2
data.ts.boxcox = (data.ts^lambda-1) / lambda 
fig_nums("test-normality", paste('Test Normality of ', default_ylab, sep = '')) %>% cat()
doTestNormality(data.ts.boxcox)
```

The Box-Cox transformation **did** help to improve the normality of the series because:
- the dots are moreso aligned with the red line in the QQ plot
- However, it isn't a perfect fit with QQ between ~(-2:0.5) does not fitting the QQ plot line
- The Shapiro test p-value(0.5463) > 0.05
- Possible intonation point or bimodal data in the data

## Diff 
### Without Box-Cox

```{r diff-transforms}
doDiffAndPlot(data.ts, 1, F, F) # p = 0.3601 no lags, p is insignificant 
doDiffAndPlot(data.ts, 2, F) # p = 0.1643 lag = 4 p is insignificant
doDiffAndPlot(data.ts, 3, F, F) # p = 0.3768 lag = 1, 4 = p is significant
```

- Our third diff doesn't add any further value
- Best is second diff with p 0.1643
- EACF We can't see any meaningful shelf, which indicates white noise series and possible non-stationarity

## Diff 
### With Box-Cox

```{r}
fig_nums('boxCox diff 0', 'BoxCox without any diff') %>% cat()
doDiffAndPlot(data.ts.boxcox, 0, T) # p < 0.05 = 0.01941, lag = 1 but it still has trend
doDiffAndPlot(data.ts.boxcox, 1, F) # p is higher 0.3282, better trend! Looks stationary
doDiffAndPlot(data.ts.boxcox, 2, F) # p is better 0.09166 pacf lag of 1 @4 pacf
doDiffAndPlot(data.ts.boxcox, 3, F) # p is better 0.14 pacf lag of 1 @4 pacf
```

```{r}
fig_nums('boxCox diff 4', 'BoxCox with diff of 4') %>% cat()
doDiffAndPlot(data.ts.boxcox, 4, T, F) # p is better 0.02 pacf lag of 1 @4 pacf
```

- Box-Cox transform has significance `p-value < 0.05 = 0.01941`
- Box-Cox transform with 4 differences has significance `p-value < 0.05 = 0.01997`
- Box-Cox with no difference still shows a trend
**Therefore we select Box-Cox with a diff(4)**

```{r}
# set variable for model testing:
data.ts.boxcox4 <-diff(data.ts.boxcox, difference = 4)
fig_nums('Normal Q-Q Plot test', 'Normal Q-Q Plot test') %>% cat()
doTestNormality(data.ts.boxcox4)
```

The Box-Cox transformation **did** help to improve the normality of the series because:
- the dots are more aligned with the red line in the QQ plot (than solely with the Box-Cox transformation)
- the Shapiro test p-value(0.5948) > 0.05

From ACF and PACF:
`{arima(p,d,q)} = {arima(1,4,1)}`

```{r}
fig_nums('Eacf test', 'Eacf test') %>% cat()
eacf(data.ts.boxcox, ar.max=3, ma.max=3)
```

From the eacf plot we can select: 
- `{arima(0,4,1)}`
- `{arima(1,4,0)}`

We can also select the following values which are close to the shelf:
- `{arima(1,4,1)}`
- `{arima(2,4,0)}`

## Residuals: `BIC` Table

```{r warning=FALSE}
fig_nums('Residual BIC Table', 'Residual BIC table') %>% cat()
par(mfrow=c(1,1))
armasubsets(y=data.ts,nar=3,nma=3,y.name='test',ar.method='ols') %>% plot()
```

From the `BIC` residual plot, we can also extract the models:
`{arima(1,4,2)}, {arima(1,4,1)}, {arima(0,4,2)}, {arima(0,4,1)}` 
Note: arima(0,4,1) is duplicated in eacf plot

The final set of possible models `{arima(p,d,q)}` is:
`{arima(1,4,1), arima(0,4,1), arima(1,4,0), arima(1,4,2)}, {arima(0,4,2)}, {arima(2,4,0)}`

## Arima Model

```{r}
getModelCoef <- function(pdq) {
  cat('\nmodel: arima(', pdq, ')\n')
  methods=c('CSS','ML')
  for (i in methods) {
    cat(i, '\n')
    model = arima(data.ts,order=pdq,method=i)
    coef = coeftest(model)
    modelName <- paste("model", i, sep = "")
    modelToScore <- assign(modelName, model)
    totalResultLines <- pdq[1] + pdq[3]
    startResult <- totalResultLines*3# because coeftest() returns a s4 object
    pValues <- 1:totalResultLines %>%
      map(~ {coeftest(model)[(.x + startResult)] %>% round(6) %>% paste()})
    isPSignifcant <- function(p) { 
      ifelse(p < 0.05, 'p < 0.05 significant', 'p > 0.05 insignificant')
    }
    pValues %>% rbind(isPSignifcant(pValues), '\n') %>% paste() %>% cat()
  }
}
cat('Arima Model coeff p-values (rounded to 6 decimal places):\n')
model_041_p <- getModelCoef(pdq=c(0,4,1))
model_042_p <- getModelCoef(pdq=c(0,4,2))
model_140_p <- getModelCoef(pdq=c(1,4,0))
model_141_p <- getModelCoef(pdq=c(1,4,1))
model_142_p <- getModelCoef(pdq=c(1,4,2))
model_240_p <- getModelCoef(pdq=c(2,4,0))
model_242_p <- getModelCoef(pdq=c(2,4,1))
```

- arima(1,4,2), arima(2,4,1), arima(2,4,0) show insignificant p-values > 0.05, 
- These are no longer models we will consider
- The remaining models are: `model_041, model_042, model_140, model_141`

```{r}
model_041_ml = arima(data.ts,order=c(0,4,1),method='ML') # ARIMA(0,4,1)
model_042_ml = arima(data.ts,order=c(0,4,2),method='ML') # ARIMA(0,4,2)
model_140_ml = arima(data.ts,order=c(1,4,0),method='ML') # ARIMA(1,4,0)
model_141_ml = arima(data.ts,order=c(1,4,1),method='ML') # ARIMA(1,4,1)

# AIC and BIC values
sort.score(AIC(
  model_041_ml, model_042_ml, model_140_ml, model_141_ml ),
  score = "aic")
sort.score(BIC(
  model_041_ml, model_042_ml, model_140_ml, model_141_ml ),
  score = "bic")
```

- arimia(0,4,2) ranks highest according to he `sort.score` formula in both AIC and BIC

```{r}
# arima(0,4,2) entire coeficient's output for for completeness:
arima(data.ts,order=c(0,4,2),method='ML') %>% coeftest()
```

## Residuals

```{r warning=FALSE}
residual.analysis <- function(model, std = TRUE){
  if (std == TRUE){
    res.model = rstandard(model)
  }else{
    res.model = residuals(model)
  }
  par(mfrow=c(3,2),
      oma = c(1,1,0,0) + 0.1,
      mar = c(2,2,2,2) + 0.1)
  plot(
    res.model,
    type='o',
    xlab='years',
    ylab='Standardised residuals',
    main="Time series plot of standardised residuals")
  abline(h=0)
  hist(res.model,main="Histogram of standardised residuals")
  qqnorm(res.model,main="QQ plot of standardised residuals")
  qqline(res.model, col = 2)
  acf(res.model,main="ACF of standardised residuals")
  print(shapiro.test(res.model))
  k=0
  LBQPlot(res.model, lag.max = length(model$residuals)-1 , StartLag = k + 1, k = 0, SquaredQ = FALSE)
  #resPlotsTitle <- paste('Figure', figureCount, '. Standardised residuals of ', default_ylab, sep = '')
  #title(resPlotsTitle, side = 3, line = -33, outer = TRUE, cex.adj = 3)
  #figureCount <- figureCount + 1
  par(mfrow=c(1,1))
}
fig_nums('Residual Analysis', 'Residual Analysis') %>% cat()
```

### Residuals for arima(0,4,1)

```{r warning=FALSE}
residual.analysis(model = model_041_ml)
```

- Time-series standardised residuals: cannot observe a trend
- histogram: data is not normal and left skewed
- Ljung-Box test: One dots below red line
- Q-Q plot: tailed, doesn't fit 
- ACF: lag of 1, indicates we have not modelled all of the data
- Shapiro-Wilk: significant

**ACF and Ljung-Box shows a timeseries patern not captured in this model**

### Residuals for arima(0,4,1)

```{r warning=FALSE}
residual.analysis(model = model_041_ml)
```


- Time-series standardised residuals: cannot observe a trend
- histogram: data is not normal and left skewed, possibly bimodal
- Ljung-Box test: One dots below red line
- Q-Q plot: tailed, doesn't fit 
- ACF: lag of 1, indicates we have not modelled all of the data
- Shapiro-Wilk: significant

**ACF and Ljung-Box shows a timeseries patern not captured in this model**

### Residuals for arima(0,4,2)

```{r warning=FALSE}
residual.analysis(model = model_042_ml)
```


- Time-series standardised residuals: cannot observe a trend
- histogram: data is not normal and left skewed
- Ljung-Box test: all dots > red line
- Q-Q plot: left-tailed, doesn't fit 
- ACF: no lags, indicating white noise
- Shapiro-Wilk: significant

**We cannot find any meaningful timeseries observations from the residuals, indicating a good model fit**

### Residuals for arima(1,4,0)

```{r warning=FALSE}
residual.analysis(model = model_140_ml)
```

- Time-series standardised residuals: cannot observe a trend
- histogram: data is not normal and left skewed
- Ljung-Box test: all dots > red line
- Q-Q plot: left-tailed, doesn't fit 
- ACF: no lags, indicating white noise
- Shapiro-Wilk: significant

**We cannot find any meaningful timeseries observations from the residuals, indicating a good model fit**

### Residuals for arima(1,4,1)

```{r warning=FALSE}
residual.analysis(model = model_141_ml)
```

- Time-series standardised residuals: cannot observe a trend
- histogram: data is not normal and left skewed
- Ljung-Box test: one dot on red line
- Q-Q plot: left-tailed, doesn't fit 
- ACF: Possible lag at 5, may not be significant
- Shapiro-Wilk: insignificant

**ACF and Shapiro-Wilk shows a timeseries patern not captured in this model**

### Residual analysis conclusion

Residuals are not apparent in `arima(1,4,0)` and `arima(0,4,2)`

## Forecast

```{r message=FALSE}
library(forecast)
figureName <- fig_nums('Forecast of growth', 'Forecast of growth') %>% cat()
# lambda of 0.35 is from Box-Cox transformation:
fit = Arima(data.ts.raw,c(0,4,2), lambda = lambda) 
# hack to hide negative values:
ylim <- c(500, 14000) 
plot(
  forecast(fit,h=10), 
   ylim=ylim, 
   ylab=figureName
   )
```

- The forecast shows an almost exponential positive trend in growth, with quite a wide confidence interval

NOTE: negative trend has been hidden on this plot, because it's not possible for fish to lay a negative number of eggs. This is a plot side-effect generated because lambda doesn't equal 0

## Conclusion

The results of the above tests summarised:

- `p-value(0.02) < 0.05` in our adfTest
- `ma2 and ma1 p < 0.05` in coefficient test
- Residuals are not apparent in `arima(1,4,0)` and `arima(0,4,2)`

The AIC and BIC score for is higher for `arima(0,4,2)`, which is why we have
selected `{arima(0,4,2)}` as our best fitting time-series model

Please note that the dataset is quite small, so we have limited most of our 
decisions to programatic output

The forcast shows a possible 3,000,000,000 eggs laid in 2006. 
Given the dataset is limited to 16 values, a more detailed dataset should be examined to check accuracy of fit. The model may also be adjusted because of the [scarcity of food and high mortality rate of larval Bloaters](https://en.wikipedia.org/wiki/Coregonus_hoyi)

If the model still fits, recommendations include investing in the cavier industry in Lake Huron
