---
title: "Water tasting week 2"
output: html_notebook
---

From here: https://docs.google.com/document/d/1WzaKKXAQiUY0mMSSeXoG_Vayt603L2dWFCOjeCWmugE/edit

TODO:
1.1 Import the data into RStudio and assign the appropriate labels to the following factors: Gender, Class, UsuallyDrink, First, Second, Third, Fourth.
1.2 Explore the first, second, third and fourth preferences given by students. Is there evidence of preference for a specific brand?
1.3 Explore the Preference variable. What is the modal preference?
1.4 Does First preference depend on gender?

# 1.1 Import the data into RStudio and assign the appropriate labels to the following factors:
```{r}

setwd("~/Documents/code/tldr/data-science/stats-math1324/2week/data")
library(readr)
Water_taste_Water_taste <- read_csv("Water_taste - Water_taste.csv")
# View(Water_taste_Water_taste)
head(Water_taste_Water_taste)

```

# assign the appropriate labels to the following factors: Gender, Class, UsuallyDrink, First, Second, Third, Fourth.

```{r}

setwd('./data/')
library(readr)
Water_taste_Water_taste <- read_csv("Water_taste - Water_taste.csv")
# head(Water_taste_Water_taste)

Water_taste_Water_taste$Gender

Water_taste_Water_taste$Gender <- factor(Water_taste_Water_taste$Gender, 
  levels = c("1","2"),
  labels = c("Female","Male")
  )

Water_taste_Water_taste$Class <- factor(Water_taste_Water_taste$Class, 
  levels = c("F", "SO", "J", "SR", "O"),
  labels = c("freshman", "sophomore", "junior", "senior", "other")
  )

Water_taste_Water_taste$UsuallyDrink <- factor(Water_taste_Water_taste$UsuallyDrink, 
  levels = c("B", "F", "T", "NA"),
  labels = c("bottled", "filtered", "tap", "not applicable")
  )

Water_taste_Water_taste$First <- factor(Water_taste_Water_taste$First, 
  levels = c("A", "B", "C", "D"),
  labels = c("Fiji", "Aquafina", "Sam’s Choice", "none")
  )

Water_taste_Water_taste$Second <- factor(Water_taste_Water_taste$Second, 
  levels = c("A", "B", "C", "D"),
  labels = c("Fiji", "Aquafina", "Sam’s Choice", "none")
  )

Water_taste_Water_taste$Third <- factor(Water_taste_Water_taste$Third, 
  levels = c("A", "B", "C", "D"),
  labels = c("Fiji", "Aquafina", "Sam’s Choice", "none")
  )

Water_taste_Water_taste$Fourth <- factor(Water_taste_Water_taste$Fourth, 
  levels = c("A", "B", "C", "D"),
  labels = c("Fiji", "Aquafina", "Sam’s Choice", "none")
  )

head(Water_taste_Water_taste)


```

cheatsheet:
mydata$var1 <- factor(mydata$var1, 	#Assign labels to “var1” from data 
  levels = c("1","2"),			#object “mydata”
  labels = c("Female","Male"))		

