---
output:
  html_document:
    fig_height: 4
    fig_width: 8
    highlight: kate
    theme: paper
subtitle: Assignment 3 - Semester 1, 2019
author: "Phil Steinke s3725547 Ash Olney s3686808"
title: "MATH1318 Time Series"
---

## Executive Summary

TODO:
This report examines the .

### Goals:

HEADINGS INDENTED BELOW

- [ ] In short, the challenge is to find the best fitting model 
to the given cryptocurrency series.
- [ ] captions 
`fig_nums(
  'initial ts data', 
  'initial time-series data with no transformation') %>% cat()`

Your task is to:

- [ ] analyse the data by using the analysis methods covered in MATH1318 
Time Series Analysis course in this semester
- [ ] accurately predict the value of bitcoin for the next 10 days, 
- [ ] and prepare a comprehensive analysis report including 
  - [ ] descriptive analysis 
  - [ ] proper visualisation 
  - [ ] model specification 
  - [ ] model fitting and selection 
  - [ ] and diagnostic checking
- [ ] You will include a mean absolute scaled error (MASE), 
for each of model fits and forecasts. 
- [ ] You’ll use the real values of daily bitcoin for 10 days of the forecast 
period (_25th of February_ - _6th of March 2019_) to compute `MASE` for 
forecasts using the R function here.
- [ ] present your results in a written report format and as an oral 
presentation.
- [ ] QQ plot

## R code 15%

```{r load-packages-and-scripts, message=FALSE, warning=FALSE}
# devtools::install_git('https://gitlab.com/botbotdotdotcom/packagr')
library(packagr)
packages <- c(
  'AID','CombMSC','fGarch','forecast','FSAdata','fUnitRoots',
  'lmtest','magrittr','nortest','rugarch','TSA','captioner')
packagr(packages) # alpha package to check, install and load packages
# setwd('./Users/phil/code/data-science/uni/time-series/assignment03')

# current_path = rstudioapi::getActiveDocumentContext()$path
# setwd(dirname(current_path ))
# print( getwd() )
source('/Users/phil/code/data-science/uni/time-series/assignment03/TSHandy.r') 
# check if updated
source('/Users/phil/code/data-science/uni/time-series/assignment03/utils.r') 
# moved all functions to here
source('/Users/phil/code/data-science/uni/time-series/assignment03/MASE.r')
doDiffAndPlot <- doDiffAndPlot # fix lint no symbol named 'X' in scope
```

```{r, env-variables}
startDate = '2013-04-27' #yyyy-mm-dd
endDate = '2019-02-24'
default_ylab = 'Bitcoin EOD closing price US$'
default_xlab = 'Year'
```

## The Data

```{r}
data <-
  read.csv("./Bitcoin_Historical_Price.csv",
  sep = ",",
  fill = TRUE)
# data %>% class() # data.frame needs to be converted to time series
```

```{r, data-wrangle}
# remove commas from currency - credit to Zoe
data$Close <- as.numeric(as.character(gsub(",","",data$Close)))

inds <- seq(as.Date(startDate), as.Date(endDate), by = "day")
data.ts <- ts(
  as.vector(data[,2]),
  start = c(2013, as.numeric(format(inds[1], "%j"))),
  frequency = 365.25) # or 365?
data.ts.raw <- data.ts
# Code source: https://stackoverflow.com/a/33129922
```

```{r, data-embed}
# Embed the data
readLines("./Bitcoin_Historical_Price.csv") %>%
  paste0(collapse = "\n") %>%
    openssl::base64_encode() -> encoded
```

### Source

The dataset you will focus on has gathered from [coinmarketcap.com] includes 
the daily closing price of bitcoin from the _27th of April 2013_ to the 
_24th of February 2019_.

[Download BitCoin CSV data](`r sprintf('data:text/csv;base64,%s', encoded)`)

### Historical Fluctuation

The value of bitcoin is determined by what people are willing to pay for it, 
and is very volatile, fluctuating wildly from day to day. 
- In April 2013, the value of 1 bitcoin (BTC) was around $100 USD. 
- At the beginning of 2017 its value was $1,022 USD and 
- by the 15th of December, it was worth $19,497. 
- As of the 3rd of March 2018, 1 BTC was sold for $11,513 USD. 
- But now it came back to the window of $3800USD - $4000USD.

## Trend Models

```{r initial-data}
doPar(mfrow = c(1, 1))
plot(
  data.ts,
  type = 'l',
  main = fig_nums("bitcoin_ts_data", 
                  "Bitcoin time series data"),
  ylab = default_ylab)
```

- There is no overall trend
  <!-- - There was positive trend until late 2017 -->
- No obvious seasonality
- There looks like there could be an intervention point after the start of 2017
- It is evident from the plot that there are changes in variance throughout 
the series
- There are clusters of high and low variance
 
```{r}
y = data.ts        # Read data into y
x = zlag(data.ts)  # Generate first lag
index = 2:length(x) # Create an index to get rid of the first NA value
cor(y[index], x[index])
doPar(mfrow = c(1, 1))
plot(
  y = data.ts,
  x = x,
  ylab = 'Closing Price',
  xlab = 'Previous Day Closing Price',
  main = fig_nums("Scatter_neighbouring", 
                  "Scatter plot of neighbouring Closing Prices")
)
```

- The high correlation of the data with the first lag shows that successive 
data points are related

### Linear trend model

```{r, linear-trend-model}
model_linear = lm(data.ts ~ time(data.ts))
summary(model_linear)
```

- The linear model has a positive trend
- The R-Squared value of 0.47 is relatively low
- This model only explains 47% of variation in the data

```{r}
doPar(mfrow = c(1, 1))
plot(
  data.ts,
  main = fig_nums('linear_model_abline', 'Bitcoin Data with Linear Model'),
  ylab = default_ylab)
abline(model_linear)
```

```{r message=FALSE, warning=FALSE}
residual.analysis(
  model_linear, 
  std = TRUE, 
  start = 1,   
  title = fig_nums('linear_residual', 'Linear model residual analysis')
)

# TODO: MASE for linear model
```

- The standardised residuals are not white noise and show that the model has 
not captured the trend of the data
- histogram:  left skewed
- Ljung-Box test: all lags are significant
- Q-Q plot: left-tailed, doesn't fit along one line, therefore not normally 
distributed
- ACF: there are significant lags which does not fit with a white noise series
- Shapiro-Wilk: significant, therefore reject the null that the residuals are 
normally distributed
- TODO: MASE value for this model

## Quadratic Model

```{r, quadratic-model}
t = time(data.ts)
t2 = t ^ 2
model_quadratic = lm(data.ts ~ t + t2) # label the quadratic trend model
summary(model_quadratic)
```

```{r}
doPar(mfrow = c(1, 1))
plot(
  ts(fitted(model_quadratic)),
  ylim = c(min(c(
    fitted(model_quadratic), as.vector(data.ts)
  )),
  max(c(
    fitted(model_quadratic), as.vector(data.ts)
  ))),
  ylab = default_ylab,
  main = fig_nums(
    "quadratic_model",
    "Bitcoin Data with Quadratic Model")
)
lines(as.vector(data.ts))
```

```{r}
residual.analysis(
  model_quadratic, std = TRUE, start = 1,
  title = fig_nums(
    'quadratic_model_residual', 
    'Quadratic model residual analysis'))

# TODO: MASE for quadratic model
```

- The residuals are largely unchanged from the linear plot
- The standardised residuals are not white noise and show that the model has not
captured the trend of the data
- histogram:  left skewed
- Ljung-Box test: all lags are significant
- Q-Q plot: left-tailed, doesn't fit along one line, therefore not normally 
distributed
- ACF: there are significant lags which does not fit with a white noise series
- Shapiro-Wilk: significant, therefore reject the null that the residuals are 
normally distributed
- TODO: MASE value for this model

```{r, plot-no-diff}
doDiffAndPlot(data.ts, 0, 'Bitcoin with no diff')
```
- acf plot slowly decaying lags i.e. not stationary
- the adf test confirms that the series is not stationary 
- the series will be transformed and/or differenced before fitting ARIMA

```{r, message=FALSE, warning=FALSE}
# TODO: look at code to see where he uses it 
# I think we should scrap this? - ash
# bc.search = BoxCoxSearch(y = data.ts)
```

```{r, plot-first-diff}
# NEW: testing diff without log 
doDiffAndPlot(data.ts, 1, 'BitCoin first diff')
```

- appears to be volatility clustering in plot
- Therefore: we will transform the data with log or Box Cox

### Confidence interval of Lambda

```{r, confidence-interval, warning=FALSE}
# check the confidence interval of lambda
doPar(mfrow = c(1,1))
boxcoxCi <- BoxCox.ar(data.ts, method = "yule-walker")$ci
boxcoxCi
title(fig_nums('boxcox_ci','Confidence Interval of Lambda of Bitcoin'), line = -1.5, outer = TRUE)
```

lambda == 0

- The confidence interval does include 0, so we _will_ do a log transform

## Log Transform

```{r, log_transform}
data.ts_log = log(data.ts)
doDiffAndPlot(
  data.ts_log, 0, T, T,
  fig_nums("log_transform", "Log transform"))
```

- the adf test p-value is not significant meaning the series is not stationary
- acf still has the same pattern

### Log first diff

```{r, log-first-diff}
doDiffAndPlot(
  data.ts_log, 1, T, T,
  fig_nums('log_first_diff', 'Log first diff'))
data.ts_log_diff1 = data.ts_log %>% diff(differences = 1)
```

- the adf test p-value is significant - therefore the series is now stationary
- However, doesn't have constant variance - an assumption for ARIMA models
- there are many significant serial correlations in the data which suggests 
that the data exhibits serial dependence

### Log Second diff

```{r, log-second-diff}
doDiffAndPlot(data.ts_log, 2, T, T,
  plotTitle = fig_nums('log_second_diff', 'Log second diff'))
```

- single AR lag 
- lots of diffs in MA
- don't need, overdifferencing
- aka overPhil

### M TODO:

> ACF, PACF and EACF all shows pattern of white noise for the correlation 
> structure. However, there is an ARCH effect present in the series.

```{r}
doPar(mfrow = c(1, 1))
McLeod.Li.test(
  y = data.ts_log_diff1,
  main = fig_nums(
    "mcleod_first_diff_log",
    "McLeod-Li Test Statistics for First Difference of Log Bitcoin"))
```
- McLeod-Li test is significant at 5% level of significance for all lags. 
- This gives a strong idea about existence of volatility clustering.

```{r}
doPar(mfrow = c(1, 1))
qqnorm(
  data.ts_log_diff1,
  main = fig_nums(
    "qq_first_diff_log",
    "Q-Q Normal Plot of First Difference of Log Bitcoin")
  )
qqline(data.ts_log_diff1)
```
- Shows volatility clustering

```{r absolute_transformation}
abs.data.ts_log_diff1 = abs(data.ts_log_diff1)
doDiffAndPlot(
  abs.data.ts_log_diff1, 
  0, T, T, 
  fig_nums("absolute_transformation", "Absolute Transformation"))
```

NOTE: TODO: After the absolute value transformation, we observe many 
signficicant lags in both ACF and PACF. Also, EACF does not suggest an 
ARMA(0,0) model.
From the EACF, we can identify ARMA(1,1) and ARMA(1,2) models for absolute 
value series.
These models correspond to parameter settings of [max(1,1),1] and 
[max(1,2),1]. So the corresponding
tentative GARCH models are GARCH(1,1) and GARCH(2,1)

```{r}
sq.data.ts_log_diff1 = data.ts_log_diff1 ^ 2
doDiffAndPlot(
  sq.data.ts_log_diff1, 0, T, T, 
  fig_nums('square-transform','Square transformation'))
```

NOTE: TODO:
After the square transformation, we boserve many signficicant lags in both 
ACF and PACF. Also, EACF does not suggest an ARMA(0,0) model.
From the EACF, we can identify ARMA(1,1) and ARMA(1,2) models for squared 
series. 
These models correspond to parameter settings of [max(1,1),1] and 
[max(1,2),1]. So the corresponding 
tentative GARCH models are GARCH(1,1) and GARCH(2,1).

### Garch

```{r, garch}
# ash says: I think this is fitting the garch model
model_s_garch <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(1, 0), include.mean = FALSE),
  distribution.model = "norm"
)
# m.2 <- ugarchfit(spec = model_s_garch, data = data.ts_log_diff1)
ugarchfit(spec = model_s_garch, data = data.ts_log_diff1)
```

## ARIMA MODELLING 

```{r, arima}
# doDiffAndPlot(data.ts, 0)
```

```{r}
adfTest(data.ts)
# doDiffAndPlot(data.ts, 1)
```

### TODO: Arima subset

```{r message=FALSE, warning=FALSE}
res = armasubsets(
  y = data.ts,
  nar = 5,
  nma = 5,
  y.name = 'test',
  ar.method = 'ols')
doPar(mai=c(1,0.5,0.5,0.5))
plot(res)
title(
  fig_nums('arima-subset', 'Arima Subset'), 
  line = -2, 
  outer = TRUE)
```

```{r}
# FIXME:error here
# TODO: add other models
# can't add diff.data.ts
# AIC and BIC values
# sort.score(AIC(
  # model_041_ml, model_042_ml, model_140_ml, model_141_ml ),
  # score = "aic")
# sort.score(BIC(
  # model_041_ml, model_042_ml, model_140_ml, model_141_ml ),
  # score = "bic")
```

### Residuals for arima(,,)

```{r, warning=FALSE}
# residual.analysis(model = )
```

- Time-series standardised residuals: 
- histogram: 
- Ljung-Box test: 
- Q-Q plot: 
- ACF: 
- Shapiro-Wilk: 

## Forecast

```{r, message=FALSE}
fit = Arima(data.ts.raw, c(0, 4, 2), lambda = 0)
doPar(mfrow = c(1, 1))
forecast(fit, h = 10) %>% 
  plot(
    main = fig_nums('forecast-growth', 'Forecast of bitcoin growth'),
    xlim = c(2018, 2020))
```

- TODO: interpret forecast

## TODO: Conclusion

The results of the above tests summarised:

TODO: interpret

[coinmarketcap.com]: coinmarketcap.com

```{r}
#outputting plots
plots.dir.path <-
  list.files(
     tempdir(), pattern="rs-graphics", full.names = TRUE);
plots.png.paths <-
 list.files(
   plots.dir.path, pattern=".png", full.names = TRUE)
```

```{r}
file.copy(
  from=plots.png.paths,
  to="~/code/data-science/uni/time-series/assignment03/presentation-ts-assignment03/assets")
```
