---
title: "Module 5 tasks - Specification of ARIMA models"
subtitle: time-series tasks
---

# Module 5: Tasks
## Models for Non-Stationary Time Series

```{r Setup, message=FALSE, warning=FALSE, include=FALSE, warnings=0}
rm(list = ls())
cat("\014") # clear everything
localWorkingDirectory <- "~/code/data-science/uni/time-series/05-model-specification/tasks/raw-data"
library('TSA')
library('tseries')
library('tidyr')
# set.seed(6452135)
setwd(localWorkingDirectory)
```

```{r}
# arima.data1 = arima.sim(list(order = c(2,2,1), ar = c(0.19,0.6), ma=0.1), n = 100) 
# plot(arima.data1)
# write.table(arima.data1,file="data.sim.csv",row.names = FALSE,col.names=FALSE)
```

```{r}
### 1. Gold
The dataset “gold” contains the daily price of gold (in dollars per troy ounce) for the 252 trading days of year 2005. Load the dataset and do the following tasks:

#--- Task 1 ---
data(gold) # Load data. Here no need to convert it to ts object because
# it is already ts object. See below:
class(gold)
gold
par(mfrow=c(1,1)) # ~old_todo~ in r studio - what is par?
# ?par
plot(gold,type='o',ylab="Rainfall", main='Los Angelos rainfall series')
# In the time series plot there is trend, no seasonality, no obvious intervention, and no changing 
# variance and succeeding observations imply the existence of autoregressive 
# behavior.


par(mfrow=c(1,2))
acf(gold) #Trend is apparent from ACF and PACf plots
pacf(gold)
par(mfrow=c(1,1))
# Slowly decaying pattern in ACF and very high first correlation in PACF
# implies the existence of trend and nonstationarity.

# Apply ADF test with default settings
adf.test(gold)
# With a p-value of 0.6665, we cannot reject the null hypothesis stating that
# the series is non-stationary.

# Let's first apply the box-Cox transformation.
gold.transform = BoxCox.ar(gold)
gold.transform$ci
# Mid-point of interval is -1.8. So we will take lambda as -1.8
lambda = -1.8
BC.gold = (gold^lambda-1)/lambda
qqnorm(BC.gold)
qqline(BC.gold, col = 2)
shapiro.test(BC.gold)
# The Box-Cox transformation did not help to improve normality of the series 
# because the dots are not aligned with the red line in QQ plot and 
# p-value of the Shapiro test is less than 0.05.

#Let's calculate the first difference and plot the first differenced series
diff.BC.gold = diff(BC.gold)
plot(diff.BC.gold,type='o',ylab='Quarterly earnings ')
# After applying the the first difference the series became detrended and stationary.
# Now there is some changing varinace come the light.

# Let's apply ADf unitroot test to the differenced series.
adf.test(diff.BC.gold)
# With a p-value of 0.01, we reject the null hypothesis stating that
# the series is non-stationary; hence, we conclude that the first differencing 
# make the series staionary.

# Specify the orders using ACf, PACF, EACF and BIC over the differenced series
par(mfrow=c(1,2))
acf(diff.BC.gold)
pacf(diff.BC.gold)
par(mfrow=c(1,1))
# Because there is no significant lags seen in ACF and PACF it is not possible to 
# proceed over the ACF and PACF for this series. The patterns in ACF and PACF
# implies existence of white noise behaviour.

# diff.BC.gold
eacf(diff.BC.gold)
# In EACF, the vertex includes the orders of p=0 and q=0 in accordance with the 
# inference drawn from the ACF and PACF. But from EACF, we can include the models
# with p=0; q=1 and p=1; q=0. Coreespondingly, the set of candidate models is
# {ARIMA(1,1,0), ARIMA(0,1,1)}. Here the "1" in the middle comes from the 
# first differencing.

# Let's display the BIC table
res = armasubsets(y=diff.BC.gold,nar=14,nma=14,y.name='test',ar.method='ols')
plot(res)
# In the BIC table shaded columns correspond to AR(1) and AR(8) coefficients and 
# there is one MA effect(lag 8) supported by BICs. Since lag 8 is too large and,
# we observed white noise from ACF and PACF, we can include ARIMA(1,1,0)
# and ARIMA(8,1,0) models.

# In conclusion the set of candidate models is {{ARIMA(1,1,0), ARIMA(0,1,1)}}. 
# In the next modules we will fit these models and
# select the best one according to some selection measures.
```

- Conduct a hypothesis test to decide on the existence of a trend in this series.

```{r}

```

- Overcome the nonstationary nature of this series by using suitable tools.

```{r}

```

### Specify the orders of ARIMA(p,d,q) model by using:

#### ACF and PACF plots

```{r}

```

#### EACF plot

```{r}

```

#### BIC table

```{r}

```

- Do the specification tools agree on a specific ARIMA model? Which one do you choose?

________________________________________________

### 2. JJ

_The dataset “JJ” contains quarterly earnings per share for the Johnson & Johnson Company._ 



Load the dataset and do the following tasks:
```{r}

data(JJ)# Load data. Here no need to convert it to ts object because
# it is already ts object. See below:
class(JJ)
par(mfrow=c(1,1))
plot(JJ,type='o',ylab='Earnings ', main='Quarterly earnings ')
# In the time series plot there is:
# - trend, 
# - a repeating pattern (seasonality) and 
# - # changing (increasing variance) and 
# - bouncing observations around the mean level
# imply the existence of moving average behaviour.
```

#### Conduct a hypothesis test to decide on the existence of a trend in this series.
```{r}
par(mfrow=c(1,2))
acf(JJ)
pacf(JJ)
par(mfrow=c(1,1))
# Slowly decaying pattern in ACF and very high first correlation in PACF
# implies the existence of 
# - trend and 
# - non-stationarity.
```

#### Overcome the non-stationary nature of this series by using suitable tools
```{r}
# Let's first apply the Box-Cox transformation.
jj.transform = BoxCox.ar(JJ)
jj.transform$ci
lambda = 0.2 # The mid point of the interval 
BC.jj = (JJ^lambda-1)/lambda # apply the Box-Cox transformation
par(mfrow=c(1,1))
qqnorm(BC.jj)
qqline(BC.jj, col = 2)
shapiro.test(BC.jj)
# In the QQ plot the tails of the distribution is far from the normality.
# The p-value of Shapiro test is less than 0.05; hence, we have enough 
# evidence to reject the normality hypothesis. In conclusion, the Box-Cox
# transformation did not significantly help to improve normality of the observations.
plot(BC.jj,type='o',ylab='Earnings ', main='Box-Cox transformed Quarterly earnings ')
# In theime series plot, we observe that the variation in the series is 
# decreased after applying the Box-Cox transformation. But there is still trend.
# So let's apply the first difference and see if it helps.

diff.BC.jj = diff(BC.jj)
par(mfrow=c(1,1))
plot(diff.BC.jj,type='o',ylab='Quarterly earnings ')
# Now, there is only changing variance in the series after taking the first difference.
# Let's go on with the specification of the models.
adf.test(diff.BC.jj)
# The null hypothesis of non-staionarity is rejected with a p-value of 0.01; hence,
# we conclude that the first differencing make the series staionary.

```

#### Specify the orders of `ARIMA(p,d,q)` model by using `ACF and PACF` plots
```{r}
par(mfrow=c(1,2))
acf(diff.BC.jj)
pacf(diff.BC.jj)
par(mfrow=c(1,1))
# There is a slowly decaying jagged pattern in ACF and 4 significant lags in the PACF.
# So we can consider ARI(4,1) model and those smaller than this model like ARI(1,1)
# ARI(2,1) and so on.
```

#### EACF plot
```{r}
eacf(diff.BC.jj)
# Although ther is no clear vertex in EACF, we can take the row corresponding to 
# p = 3 as the vertex and include ARIMA(3,1,2), ARIMA(3,1,3) and ARIMA(4,1,2)
# model into the set of possible models.

```

#### BIC table
```{r}
par(mfrow=c(1,1))
res2 = armasubsets(y=diff.BC.jj,nar=14,nma=14,y.name='test',ar.method='ols')
plot(res2)
# In the BIC table, shaded columns correspond to AR(4), MA(3) and MA(11) coefficients and 
# So, from here we can include ARIMA(4,1,2) and ARIMA(4,1,11) models in the set of 
# candidate model.

```
- Hadar said: expected 10 times that number of parameters


#### Do the specification tools agree on a specific ARIMA model? 
<!-- # In conclusion the set of candidate models is
# {ARIMA(3,1,2), ARIMA(3,1,3), ARIMA(4,1,2), ARIMA(4,1,2)}. Here ARIMA(4,1,11)
# is excluded because it is a very big model in terms of the number of its parameters.
# In the next modules we will fit these models and
# select the best one according to some selection measures. -->

- Which one do you choose?


____________________________________________________________

### 3. Unemployment

_The dataset given in “unemployment.csv” contains yearly averages of monthly total number of unemployed people (in thousands) in Australia between 1978 and 2016. This dataset is available through_ http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/6202.0Feb%202017?OpenDocument#Time 

Load the dataset, 
convert it into a ts object and do the following tasks:
```{r}
#--- Task 3 ---
# Read data into R
unemployment <- read.table("unemployment.csv", quote="\"", comment.char="")
# Convert to ts object
unemployment = ts(unemployment,start = 1978)
class(unemployment)
```

Conduct a hypothesis test to decide on the existence of a trend in this series.

```{r}
par(mfrow=c(1,1))
plot(unemployment,type='o',ylab='Yearly average unemployment numbers')
# In the time series plot there is trend, no seasonality and no changing 
# variance and succeeding observations imply the existence of autoregressive 
# behavior.

par(mfrow=c(1,2))
acf(unemployment)
pacf(unemployment)
par(mfrow=c(1,1))
# Slowly decaying pattern in ACF and very high first correlation in PACF
# implies the existence of trend and nonstationarity.

```

Overcome the nonstationary nature of this series by using suitable tools.

```{r}
# Let's first apply the box-Cox transformation.
par(mfrow=c(1,1))
unemployment.transform = BoxCox.ar(unemployment)
unemployment.transform$ci
lambda = 1 
# The mid point of the interval is very close to 1, which corresponds to no transformation
# So we will go on with the original series.

diff.unemployment = diff(unemployment)
par(mfrow=c(1,1))
plot(diff.unemployment,type='o',ylab='Quarterly earnings ')
# There is still some trend in ther series
adf.test(diff.unemployment)
# ADF test confirms with the p-value of 0.03375 that the series is still non-staionary at 1% level of significance.
# So we will apply the second differencing.

diff.unemployment2 = diff(unemployment, differences = 2)
par(mfrow=c(1,1))
plot(diff.unemployment2,type='o',ylab="Unemployment")
# There is no trend in ther series
adf.test(diff.unemployment2)
# ADF test confirms with the p-value of 0.01 that the series is staionary with 
# the third differencing.
```

### Specify the orders of ARIMA(p,d,q) model by using:
#### ACF and PACF plots

```{r}
par(mfrow=c(1,2))
acf(diff.unemployment2)
pacf(diff.unemployment2)
# There is one significant lag in ACF and two significant lags in PACF.
# So we can include ARIMA(2,2,1) model among the tentative models.
```
#### EACF plot

```{r}
eacf(diff.unemployment2,ar.max = 5, ma.max = 5) 
# We put these argument to limit the orders p and q at 5. Otherwise, the eacf()
# function retunrs an error and displays nothing. From the output of eacf
# we include ARIMA(0,2,1) and ARIMA(1,2,1) models in the set of
# possible models.
```
#### BIC table

```{r}
par(mfrow=c(1,1))
res3 = armasubsets(y=diff.unemployment2,nar=5,nma=5,y.name='test',ar.method='ols')
plot(res3)
# In the BIC table, shaded columns correspond to AR(2), MA(1) and MA(5)
# coefficients and from here we can include ARIMA(2,2,1), ARIMA(2,2,5) models as
# the other combinations give large models.
```

- Do the specification tools agree on a specific ARIMA model? 
- Which one do you choose?
```{r}
# In conclusion the set of candidate models is
# {ARIMA(0,2,1), ARIMA(1,2,1), ARIMA(2,2,1), ARIMA(2,2,5)}. 
# In the next modules we will fit these models and
# select the best one according to some selection measures.
```

________________________________________________

4. The dataset given in “earnings.csv” file contains average weekly earnings (AUD) in Australia on each November between 1994 and 2016. This dataset is available through http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/6302.0Nov%202016?OpenDocument#Time 

- Load the dataset, convert it into a ts object and do the following tasks:
```{r}
#--- Task 4 ---
# Read data into R
earnings <- read.table("earnings.csv", quote="\"", comment.char="")
# Convert to ts object
earnings = ts(earnings,start = 1978)
class(earnings)

par(mfrow=c(1,1))

plot(earnings,type='o',ylab='Yearly average unemployment numbers')
# In the time series plot there is trend, no seasonality, no intervention, and no changing 
# variance and succeeding observations imply the existence of autoregressive 
# behavior.

par(mfrow=c(1,2))
acf(earnings)
pacf(earnings)
# Slowly decaying pattern in ACF and very high first correlation in PACF
# implies the existence of trend and nonstationarity.


```
- Conduct a hypothesis test to decide on the existence of a trend in this series.
```{r}

```
- Overcome the nonstationary nature of this series by using suitable tools.
```{r}
# Let's first apply the box-Cox transformation.
par(mfrow=c(1,1))
earnings.transform = BoxCox.ar(earnings)
earnings.transform$ci
lambda = 0 # The interval includes 0, which corresponds to log transformation
BC.earnings = log(earnings)
par(mfrow=c(1,1))
plot(BC.earnings)
# Because there was no changin varinace, transformation had no effect on the series.
# we will go on with the original series.

diff.earnings = diff(earnings)
par(mfrow=c(1,1))
plot(diff.earnings,type='o',ylab='Quarterly earnings ')
# There is still some trend in ther series
adf.test(diff.earnings)
# ADF test confirms with the p-value of 0.7564 that the series is still non-staionary at 5% level of significance.
# So we will apply the second differencing.

diff.earnings2 = diff(earnings, differences = 2)
par(mfrow=c(1,1))
plot(diff.earnings2,type='o',ylab='Quarterly earnings ')
# There is still some trend in ther series
adf.test(diff.earnings2)
# ADF test confirms with the p-value of 0.0764 that the series is still non-staionary at 5% level of significance..
# So we will apply the third differencing.

diff.earnings3 = diff(earnings, differences = 3)
par(mfrow=c(1,1))
plot(diff.earnings3,type='o',ylab='Quarterly earnings ')
# There is no trend in ther series
adf.test(diff.earnings3)
# ADF test confirms with the p-value of 0.01 that the series is staionary with 
# the third differencing.

```

### Specify the orders of ARIMA(p,d,q) model by using:

#### ACF and PACF plots

```{r}
par(mfrow=c(1,2))
acf(diff.earnings3)
pacf(diff.earnings3)
# There is one highly significant lag in ACF and two significant lags in PACF.
# So we can include ARIMA(2,3,1) models among the tentative models.


```

#### EACF plot

```{r}
eacf(diff.earnings3,ar.max = 5, ma.max = 5) 
# We put these argument to limit the orders p and q at 5. Otherwise, the eacf()
# function retunrs an error and displays nothing. From the output of eacf
# we include ARIMA(0,3,1), ARIMA(0,3,2) and ARIMA(1,3,1) models in the set of
# possible models.
```

#### BIC table

```{r}
par(mfrow=c(1,1))
res4 = armasubsets(y=diff.earnings3,nar=5,nma=5,y.name='test',ar.method='ols')
plot(res4)
# In the BIC table, shaded coulmns correspond to AR(3), AR(5), MA(1) and MA(5)
# coefficients and from here we can include ARIMA(3,3,1), ARIMA(3,3,5) models as
# the other combinations give large models.
```

- Do the specification tools agree on a specific ARIMA model? 
- Which one do you choose?
```{r}
# In conclusion the set of candidate models is
# {ARIMA(2,3,1), ARIMA(0,3,1), ARIMA(0,3,2), ARIMA(1,3,1), 
# ARIMA(3,3,1), ARIMA(3,3,5) }. In the next modules we will fit these models and
# select the best one according to some selection measures.
```

