---
title: "MATH1324 Introduction to Statistics Assignment 2"
author: "Phil Steinke s3725547"
output: html_notebook
---

#### 1. Problem Statement:
- This paper examins the wrist girth of male and females from the dataset [Heinz G, Peterson LJ, Johnson RW, Kerk CJ. 2003. Exploring Relationships in Body Dimensions. Journal of Statistics Education 11(2).](https://ww2.amstat.org/publications/jse/v11n2/datasets.heinz.html). 
- The examined the factors are `wri.di` wrist girth (in cm) and `sex` (male or female)
- I compare the differences in body dimensions between male and females against the normal disntribution of the population of the dataset.

####  2. Load Packages:
```{r, echo = TRUE, message=FALSE, warnings=0}
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(base))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library("mosaic", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library"))
setwd("~/code/tldr/data-science/stats-math1324/assignment2")
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("latticeExtra"))

```

### 3. Data: Import the body measurements data and tidy it up

```{r}
body_measurements <- read_excel("data/bdims.csv (1).xlsx", sheet = "bdims.csv") %>% as.data.frame()
wrist_measurements <- select(body_measurements, wri.gi, sex) %>% as.data.frame()
colnames(wrist_measurements) <- c("wrist.girth", "sex")
wrist_measurements <- mutate(wrist_measurements, sex=factor(sex, labels=c("female", "male")))
wrist_measurements %>% head()
```

### 4. Summary Statistics: Calculate descriptive statistics

#### Summary of wrist measurements for all sexes
```{r}
wrist_summary <- wrist_measurements$wrist.girth
summary(wrist_summary)
summary_mean <- mean(wrist_summary) %>% print()
summary_sd <- sd(wrist_summary) %>% print()
IQR(wrist_summary) # Interquartile Range
```

#### Male Summary Statistics
```{r}
male <- subset(wrist_measurements, sex == "male")$wrist.girth
summary(male)
male_mean <- mean(male) %>% print()
male_sd <- sd(male) %>% print()
IQR(male) # Interquartile Range
```

#### Female Summary Statistics
```{r}
female <- subset(wrist_measurements, sex == "female")$wrist.girth
summary(female)
female_mean <- mean(female) %>% print()
female_sd <- sd(female) %>% print()
IQR(female) # Interquartile Range
```
### 5. Distribution Fitting:
#### For males
```{r}

plot.new()
histogram(
  x = male,
  type="density",
  main = "male wrist girth (cm) compared to normal distribution",  col=rgb(0,0,0,alpha=0.0), xlim=c(12,21), ylim = c(0,0.80),
  breaks = 40,
  xlab = "wrist girth (cm)",
  ylab = "Density Histogram\n with Normal Fit")
plotDist(
  "norm",
  mean = summary_mean,
  sd = summary_sd,
  kind= "density",
  type = "b",
  col = "red",
  lwd=2,
  add=TRUE,
  xlab = "wrist girth (cm)")
plotDist(
  "norm",  
  mean = male_mean,
  sd = male_sd,
  kind= "density",
  type = "b",
  col = "blue",
  lwd=2,
  add=TRUE,
  xlab = "wrist girth (cm)")
legend("topright", 
       c("normal distribution", "male normal distribution", "density"), 
       density = c(100, 100, 100),
       fill=c("red","blue","white"))
```

### For females
```{r}
plot.new()
histogram(
  x = female,
  type="density",
  main = "female wrist girth (cm) compared to normal distribution",
  col=rgb(0,0,0,alpha=0.0), xlim=c(12,21), ylim = c(0,0.80),
  breaks = 40)
plotDist(
  "norm",  
  mean = summary_mean,
  sd = summary_sd,
  kind= "density",
  type = "b",
  col = "red",
  lwd=2,
  add=TRUE)
plotDist(
  "norm",  
  mean = female_mean,
  sd = female_sd,
  kind= "density",
  type = "b",
  col = "pink",
  lwd=2,
  add=TRUE)
legend("topright", 
       c("normal distribution", "female normal distribution", "density"), 
       density = c(100, 100, 100),
       fill=c("red","pink","white"))

```

### Interpretation:

#### What insight has been gained from the investigation?
- The mean of males is `~17.2cm` and females are `~15cm`, and the median of males is `~17.1cm` and females is `15cm`. The girth of male wrists are larger than female wrists by an average of `2cm`

#### Discuss the extent to how your theoretical normal distribution fits the empirical data
- The data fits a similar distribution 
- There are very few female cases above `17cm`
- Male, female and both sexes fit a normal bell curve

#### Make recommendations regarding the modelling of this body measurement
- Further analysis to check against other distributions
- Check against the fit with greater and fewer bins
