---
title: "Assignment 1 notes"
author: "Phil Steinke s3725547"
---
+++++++++++++++++++++++++++

Final ~old_todo~~old_later~
link to github
remove commented code
copy paste variables
check graph x and y


+++++++++++++++++++++++++++

~old_todo~~old_later~
- What are the elements of a suitable and successful data analysis
- how you might go about presenting your results in a written report format. 

## ~old_todo~~old_later~ INCLUDE? Compare to previous year
```{r}
# doTimeSeriesPlot(
#   x = zlag(ozone.ts), # Generate first lag of the series
#   xlab = 'zlag',
#   y = ozone.ts, # Read the data into y
#   ylab = 'ozone.ts',
#   title = "Compare ozone to previous/neighbouring year",
#   type='p')
# linearTrendModel_previousYear = 
#   lm(ozone.ts~zlag(ozone.ts))
# abline(
#   linearTrendModel_previousYear,
#   col = colourPallete[1])
```
~old_todo~~old_later~ where's this? Do I need it?
```{r}
# index = 2:length(x)    # Create an index to get rid of the first NA value in x
# cor(y[index],x[index]) # Calculate correlation between numerical values in x and y
```





> The residual analysis is one of the essential stages of time series modelling. So, it's good to keep in mind that you should always decide on the suitability of a model for forecasting considering the model fit (

R-square, significance tests, etc.
) along with the results of the residual analysis.



```{r}
# TRY AGAIN
data(ozone.ts)
har. = harmonic(ozone.ts,-1000) # calculate cos(2*pi*t) and sin(2*pi*t)
model4 = lm(ozone.ts~har.) # remove -1 to include the intercept term in the model
summary(model4)

plot(
  type='l',
  main="Fitted model to average monthly temperature series",
  ylab='Temperature',
  ts(
    fitted(model4),
    freq=7,
    start=c(1927)
  ),
  # ylim= range(
  #         c(
  #           fitted(model4),
  #           ozone.ts)
  #         ),
  ylim = c(
    min(c(
      fitted(model4),
      as.vector(ozone.ts)
    )), 
    max(c(
      fitted(model4), 
      as.vector(ozone.ts)
  ))),
) # ylim ensures that the y axis range fits the raw data and the fitted values
points(ozone.ts) 

```


```{r}
ozone_harmonics = 
  harmonic(ozone.ts, .5)
model_ozone_harmonics = lm(ozone.ts ~ ozone_harmonics)
summary(model_ozone_harmonics)



plot(
  col = "red",
  # lty = 100,
  main = "Fitted quadratic curve",
  type = "l",
  ylab = 'y' ,
  ts(
    fitted(model_ozone_harmonics),
    frequency = 1,
    start = 1927,
    ),
    # ylim = c(
    #   min(c(
    #     fitted(model_ozone_harmonics),
    #     as.vector(ozone.ts))), 
    # 
    #   max(c(
    #     fitted(model_ozone_harmonics),
    #     as.vector(ozone.ts)))
    # ),
    # start=c(1927),

    # ylim=
    #   range(
    #     c(
    #       fitted(
    #         model_ozone_harmonics
    #         ),
    #       ozone_harmonics
    #     )
    #   ),
)
points(as.vector(ozone.ts), type = "o")
```

```{r}
plot(
  ts(
    fitted(model4),
    freq=7,
    start=c(1927)),
    type='l',
    ylab='Temperature',
    ylim=
      range(c(
        fitted(model4),
        ozone.ts)
      ),
      main="Fitted model to average monthly temperature series"
  ) 
      # ylim ensures that the y axis range fits the raw data and the fitted values
# points(ozone.ts)
# lines(ozone.ts) 
lines(as.vector(ozone.ts), type = "o")

```

sma tta
```{r}
SMA(ozone.ts, n = 7,)

```



```{r}
beep(2)
```

```{r}
data(rwalk) # Read the data

tail(rwalk)
tail(ozone.ts)

t_ozone = time(ozone.ts)
t_random = time(rwalk)

tail(t_ozone)
tail(t_random)

length(t_ozone)
length(t_random)

model1 = lm(rwalk~t)

model1 = lm(ozone.ts~t)

print(new)

print(ozone.ts)

row.names(ozone.ts)

tail(ozone.ts)
t = time(ozone.ts) # Create time points for model fitting
model1 = lm(ozone.ts~t) # label the model as model1
h = 5 # 5 steps ahed forecasts
# Now we will implement the two-step algorithm

new = data.frame(t = c(2017:2021))

# new = data.frame(t = seq((length(t)+1), (length(t)+h), 1)) # Step 1
# Notice here that I'm using the same variable name "t" as in the 
# fitted model above, where the name of the variable showing time 
# is also "t". To run the predict() function properly,
# the names of variables in fitted model and "new" data frame
# must be the same!!!
forecasts = predict(model1, new, interval = "prediction")
# Here interval argument shows the prediction interval
print(forecasts)
```
```{r}
plot(
  ozone.ts, 
  xlim = c(1940,2022),
  ylim = c(-13,3),
  ylab = "Random walk data")
# We need to convert forecasts to time series object starting from the first 
# time steps-ahead to be able to use plot function. 
# We do this for all columns of forecasts
lines(ts(as.vector(forecasts[,1]), start = 2016), col="red", type="l")
lines(ts(as.vector(forecasts[,2]), start = 2016), col="blue", type="l")
lines(ts(as.vector(forecasts[,3]), start = 2016), col="blue", type="l")
legend("bottomleft", lty=1, pch=1, col=c("black","blue","red"), text.width = 20,
       c("Data","5% forecast limits", "Forecasts"))
```







^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^


before ash:
  
  ```{r fig.height=4, fig.width=8}
doTimeSeriesPlot <- 
  function(
    x,
    y,
    title = 'Words, mainly words'
  ) {
    
    # head(ts_data)
    head(y)
    # args <- list(...)
    # y_exists <- "y" %in% names(args)
    # if(y_exists) {
    #   y = lm(ts_data~time(ts_data))
    # } else {
    #   y = lm(ts_data~time(ts_data))
    # }
    
    # y = zlag(ts_data)
    par(bg = colourPallete[5])
    plot(
      x,
      y,
      # y = ts_data,
      # x = ts_data,
      # x = zlag(ts_data),
      type = 'o',
      ylab = '~old_todo~~old_later~ How much humans suck (annually)',
      xlab = 'year',
      col = colourPallete[9],
      pch = 19
    )
    title(
      main = title,
      col.main = colourPallete[9],
      font.main = 3)
    # points(ts_data)
  }

linearTrendModel = 
  lm(ozone.ts~time(ozone.ts))

doTimeSeriesPlot(
  y = ozone.ts,
  # y = linearTrendModel,
  x = zlag(ozone.ts),
  title = "How much humans suck (annually)"
)
linearTrendModel = 
  lm(ozone.ts~time(ozone.ts))
abline(
  linearTrendModel,
  col = colourPallete[1])

# head(ozone.ts)

```

======================================================
  
  
  <!-- ## Harmonic
  - Use the least squares approach to fit a harmonic trend to the landings time series. 
- Interpret the regression output. 
- Save the standardized residuals from the fit for further analysis. -->
  
```{r}
ozone_harmonics = harmonic(ozone.ts,)
model.ozone.har = lm(ozone.ts ~ ozone_harmonics)
summary(model.ozone.har)
plot(
  ts(fitted(model.ozone.har)),
  ylim = c(min(c(
    fitted(model.ozone.har),
    as.vector(ozone.ts)
  )), max(c(
    fitted(model.ozone.har),
    as.vector(ozone.ts)
  ))),
  ylab = 'y' ,
  main = "Fitted quadratic curve to random walk data",
  type = "l",
  lty = 2,
  col = "red"
)
lines(as.vector(ozone.ts), type = "o")
```

  
  
======================================================



- ~old_todo~~old_later~ Does it look “random?” - Probably not, tbh
- Display and interpret the time series plot for these data.
- Use least squares to fit a linear time trend to this time series. 
- Interpret the regression output. 
- Save the standardized residuals from the fit for further analysis.
- **Standardized Residuals** -  Construct and interpret the time series plot of the standardized residuals from part (b) in a new graphics window.
- **harmonic trends** Use the least squares approach to fit a harmonic trend to the landings time series. Interpret the regression output. Save the standardized residuals from the fit for further analysis.

Looking for different kinds of trends, eg season, etc.
## ~old_todo~~old_later~ Should we normalise the data?
```{r}
# ~old_todo~~old_later~ Should we normalise the data?

# dat <- data.frame(x = rnorm(10, 30, .2), y = runif(10, 3, 5))
# scaled.dat <- scale(dat)
# 
# # check that we get mean of 0 and sd of 1
# colMeans(scaled.dat)  # faster version of apply(scaled.dat, 2, mean)
# apply(scaled.dat, 2, sd)

```




# ---------------------------------------------------------------------- #








'arg' should be one of “default”, “tango”, “pygments”, “kate”, “monochrome”, “espresso”, “zenburn”, “haddock”, “textmate”

theme: paper
highlight: textmate

packages <- c(
  'knitr',
  'knit_theme'
)

opts_knit$set(out.format = "html")
knit_theme$set("baycomb")
knit_theme$get() 


# ---------------------------------------------------------------------- #

doTimeSeriesPlot <- function(ts_data, title = '') {
  plot(ts_data,
       type = 'o',
       ylab = '~old_todo~~old_later~ How much humans suck (annually)',
       xlab = 'year')
  title(main = title,
        col.main = "darkblue",
        font.main = 4)
}
titleTest = "How much humans suck (annually)"
doOzonePlot(ozone, titleTest)

doOzonePlot(ozone)

# ~old_todo~~old_later~ title refactor