---
title: Predicting Body Fat Percentage

author: "Phil Steinke s3725547"
output:
  html_notebook:
    code_folding: none
    fig_caption: yes
    highlight: espresso
    theme: paper
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---
MATH1324 Introduction to Statistics Assignment 3
#### Executive Summary
- This report investigates if there is a general, easy to determine, body circumference measurement that could be used as a general indicator for body fat percentage.
- We have the goals to: 1. Establish a formula that can convert a body circumference measurement to a predicted body fat percentage and 2. Understand how well this prediction will hold

#### Data
- **Data**: A sample of 252 men and women from was obtained from [JSE-DA]
- **Factors**: *percentage* of body fat measured using the **Brozek** method - underwater weighing technique of density. Ten other body circumference measures (e.g. abdomen) are included as factors.
- The data which was collected was then visualised to find any potential outliers and form hypothesis to test.
- This report investigates if there is a general, easy to determine, body circumference measurement that could be used as a general indicator for body fat percentage.

```{r Packages and Data import, message=F, warning=FALSE,include=FALSE, paged.print=FALSE}
packagesList <- c("readr", "base", "readxl", "dplyr", "mosaic", "gridExtra", "latticeExtra", "epitools", "qqplotr", "car", "ggplot2", "kableExtra", "mosaic", "psychometric", "psych", "Hmisc", "corrplot", "gridExtra", "grDevices")
lapply(packagesList, require, character.only = TRUE)
setwd("~/code/tldr/data-science/stats-math1324/assignment3stats")
body <- read_excel("data/Body (2) Assignment 3 data.xlsx", sheet = "Body (2) Assignment 3 data") %>% as.data.frame()
body$Sex <- factor(body$Sex, levels = c(1, 0), labels = c("Male", "Female"))
bodyCorrelation <- body %>% .[-c(1, 3:5, length(.))] # remove sex and case number
```

### 1. Test whether the mean body fat percentage for males and females are the same (two-sample t-test)
```{r}
boxplot(body$BFP_Brozek, ylab = "Outliers in Brozek score")
```
There's an outlier at `> 40` and a minimum score in `BFP_Brozek` of `0`

#### Test the assumption of normality
```{r}
 qqPlot(BFP_Brozek ~ Sex, data = body, dist = "norm")
```
There is a slight s shape to the distribution, which may warrant further investigation


#### Homogeneity Levene’s test

- H<sub>0</sub>: σ<sup>2</sup><sub>1</sub> = σ<sup>2</sup><sub>2</sub>
- H<sub>a</sub>: σ<sup>2</sup><sub>1</sub> ≠ σ<sup>2</sup><sub>2</sub>

```{r}
leveneTest(BFP_Brozek ~ Sex, data = body)
```

- H<sub>0</sub>, p > .05
- H<sub>a</sub>, p < .05

- p = 0.1395
- p > .05

The p-value for the Levene's test of equal variance for body weight between males and females was greater than `.05`
We fail to reject the null hypothesis

#### Assuming Equal Variance

- H<sub>0</sub>: μ<sub>male</sub> - μ<sub>female</sub> = 0.0
- H<sub>a</sub>: μ<sub>male</sub> - μ<sub>female</sub> ≠ 0.0

```{r}
t.test(BFP_Brozek ~ Sex, data = body, var.equal = T, alternative = "two.sided")
```
The mean of male `(18.66)` is not the same as female `(19.42)`, so fail to reject the null hypothesis

### 2. Confidence Interval
Estimate the `99% confidence interval` for the mean body fat percentage in the population.

- H<sub>0</sub>: α = 0
- H<sub>a</sub>: α ≠ 0

```{r}
confint(t.test(~ BFP_Brozek, data = body), conf.level = 0.99)
```
α is not captured `α < 18.93849 < 19.9001`
So is not captured between the upper and lower confidence interval, therefore we reject the null hypothesis.

```{r fig.height=3, fig.width=6}
favstats(~ BFP_Brozek | Sex, data = body)
histogram(~ BFP_Brozek | Sex, data = body, bins=100, nint=30, layout = c(1, 2))
```


```{r}
confint(t.test(~ BFP_Brozek, data = subset(body, subset = (Sex == "Male")), conf.level = 0.99))
```

```{r}
binom.approx(44, 109, conf.level = 0.99)
```

### 3. Researchers believe that average body fat percentage is `less than 12.5`. Test this claim.

#### One sample T Test - One tail

- H<sub>0</sub>: μ = 12.5
- H<sub>a</sub>: μ < 12.5

```{r}
t.test(~ BFP_Brozek, data = body, mu = 12.5, alternative = "less")
```

- `P = 1`
- We do not reject the null hypothesis

We could do a further test by:
- a two tailed test
- the H<sub>a</sub> being μ > 12.5
- testing a different mu based on further research
- If we test the upper tail: `t.test(~ BFP_Brozek, data=body, mu = 12.5, alternative = "greater")`
- Other tests for normal standard distribution

We could be testing for problems with the data:

- Siri and Density are also included, making the initial dataset multivariate
- Includes a minimum of `0` in `BFP_Brozek`

```{r}
favstats(~ BFP_Brozek, data = body)
filter(.data = body, BFP_Brozek == 0)
```

### 4. Regression modelling - Find the single best predictor of body fat percentage (Brozek method) using the body circumference data.

```{r name of package, echo=TRUE}
bodyCorrelationRcorrUnordered <- bodyCorrelation %>% as.matrix() %>% rcorr(type = "pearson") # Unordered bivariate
# Order our factors by Pearson correlation with with BFP_Brozek
bodyCorrelationFactorOrdered <- bodyCorrelationRcorrUnordered$r %>% as.data.frame() %>% .[1] %>% order(decreasing = T)
bodyCorrelation <- bodyCorrelation[bodyCorrelationFactorOrdered]
bodyCorrelationOrderedByPearson <- bodyCorrelation %>% as.matrix() %>% rcorr(type = "pearson") # Ordered R Correlation
#bodyCorrelationOrderedByPearson %>% colnames() # <- Pearson correlation in order
bodyCorrelationOrderedByPearson$r %>% corrplot(method = "circle", type = "upper", tl.col = "darkblue", tl.srt = 45,  p.mat = bodyCorrelationOrderedByPearson$p, sig.level = 1, insig = "blank")
pairs.panels(bodyCorrelation[1:6], method = "pearson", hist.col = "#00AFBB", density = T, ellipses = F, cor = T, ci = T, digits = 2, rug = F, breaks = 20, stars = F, show.points = F)
```
#### Write a report that explains your method for identifying the single best predictor.

Abdomen was found to be the best predictor for body fat percentage via the Brozek method,
with an `0.81*` correlation. It was found using the Pearson correlation using the `rcorr()` function.
The top 5 correlating factors are shown above on the pairplot diagram

```{r}
bodyAbdomenMaxModel <- lm(Abdomen ~ BFP_Brozek, data = body)
msummary(bodyAbdomenMaxModel)
```


- H<sub>0</sub>: ɑ = 0
- H<sub>A/sub>: ɑ ≠ 0
- P- value to test if ɑ =0.0 is <2e-16***
We reject the null hypothesis

- P-value to test if β =0.0 <2e-16 ***
- H<sub>0</sub>: β = 0
- H<sub>A/sub>: β ≠ 0
We reject the null hypothesis

#### Testing Assumption of Homoscedasticity
```{r}
mplot(bodyAbdomenMaxModel, 1)
```

```{r}
xyplot(Abdomen ~ BFP_Brozek, data = body, ylab = "Abdomen", xlab = "BFP_Brozek", panel=panel.lmbands)
```

```{r}
pf(489.9,1,250,lower.tail = FALSE ) # 7.71020869865819e-61
```
p (7.71020869865819e-61) is less than the  Confidence Level
Therefore we can reject the null hypothesis

#### Testing the Residuals
```{r}
qqPlot(bodyAbdomenMaxModel$residuals, dist="norm")
```

#### Confidence Interval

```{r}
confint(bodyAbdomenMaxModel, level = .99)
```

#### Test Intercept (α)

- H<sub>0</sub>: α = 0
- H<sub>a</sub>: α ≠ 0

The 99% Confidence Interval (CI) for α is [68.4010649, 73.832686]
H<sub>0</sub>: α = 0 is not captured between this interval, so we reject H<sub>0</sub>.

#### Testing the Slope (β)

- H<sub>0</sub>: β = 0
- H<sub>A</sub>: β ≠ 0

The 99% Confidence Interval (CI) for β is [0.9992825, 1.264792]
H<sub>0</sub>: β = 0 is not captured between this interval, so we reject H<sub>0</sub>.

```{r}
coef(summary(bodyAbdomenMaxModel))
```

#### Critique the predictive ability of the model and

The initial outlier at `> 40` and a minimum score in `BFP_Brozek` of `0` may have skewed the data

#### Draw an overall conclusion to help the investigators

- Prior to fitting the regression, a scatterplot assessing the bivariate relationship between `BFP_Brozek` and `abdomen`  was inspected.
- The scatterplot demonstrated evidence of a `positive` linear relationship.
- The overall regression model was **statistically significant**, `F(1, 250) = 489.9, p < .001`
- The results show that `Abdomen` explains 66.21% of the variability in BFP_Brozek, R2 = 0.6621. The estimated regression equation was Abdomen = 1.132037 * BFP_Brozek
- The positive slope for `abdomen` was statistically significant,
`b =1.132 , t(250) = 22.13, p < .001, 99% CI [68.4010649 ,73.832686]`.
- Final inspection of the residuals supported normality and homoscedasticity.

```{r}
save.image()
rm(list = ls())
```

[JSE-DA]: http://www.amstat.org/publications/jse/jse_data_archive.htm
