---
title: " Cars CSVModule 2 exercises week2 Phil Steinke"
output: html_notebook
---

Cars
Download the Cars.csv dataset from the Data Repository.

```{r}
library(readr)
library(forcats)
library(mosaic)

setwd("~/code/tldr/data-science/stats-math1324/02week/data/")

## Load this dataset into RStudio.
## Name the dataset Cars.

Cars <- read_csv("Cars - Cars.csv")
head(Cars)

# Ensure factors are defined correctly.

# Sports: Sports car? (1 = 'yes', 0 ='no')
Cars$Sports <- factor(Cars$Sports,
  levels = c("0","1"),
  labels = c("No","Yes")
  )
Cars$Sport_utility <- factor(Cars$Sport_utility,
  levels = c("0","1"),
  labels = c("No","Yes")
  )
Cars$Wagon <- factor(Cars$Wagon,
  levels = c("0","1"),
  labels = c("No","Yes")
  )
Cars$Minivan <- factor(Cars$Minivan,
  levels = c("0","1"),
  labels = c("No","Yes")
  )

Cars$Pickup <- factor(Cars$Pickup,
  levels = c("0","1"),
  labels = c("No","Yes")
  )
Cars$All_wheel_drive <- factor(Cars$All_wheel_drive,
  levels = c("0","1"),
  labels = c("No","Yes")
  )
Cars$Rear_wheel_drive <- factor(Cars$Rear_wheel_drive,
  levels = c("0","1"),
  labels = c("No","Yes")
  )

head(Cars, 50)
```


This dataset contains data from over 400 vehicles from 2003. The following variables, along with their coding, are included:

Vehicle_name: Model Name
Sports: Sports car? (1 = 'yes', 0 ='no')
Sport_utility: Sports utility vehicle? (1 = 'yes', 0 ='no')
Wagon: Wagon? (1 = 'yes', 0 ='no')
Minivan: Minivan?  (1 = 'yes', 0 ='no')
Pickup: Pickup?  (1 = 'yes', 0 ='no')
All_wheel_drive: All wheel drive?  (1 = 'yes', 0 ='no')
Rear_wheel_drive: Rear wheel drive?  (1 = 'yes', 0 ='no')
Retail_price: The recommended retail price ($)
Dealer_cost: The cost price for a car dealer ($)
Engine_size: Engine size in litres
Cylinders: Number of cylinders (-1 = Rotary engine)
Kilowatts: Power of the engine in kilowatts.
Economy_city: Kilometres per litre for city driving
Economy_highway: Kilometres per litre for highway driving
Weight: Curb weight of car (kg)
Wheel_base: Wheel base of car (cm)
Length: Length of car (cm)
Width: Width of car (cm)

Use this dataset to complete the following exercises.

What is the sample size?

```{r}
#tally( ~ Vehicle_name, data = Cars)
tally(~Vehicle_name, data = Cars, margins = TRUE) #Include total
# tally(~Vehicle_name, data = Cars, format="proportion", margins = TRUE) #percentage

# 428 Cars total

nrow(Cars) #428
NROW(na.omit(Cars)) #387

```


```{r}

# 2. Obtain a frequency distribution for the cylinder variable. How many cars had 4 cylinders?
freq_cylinders<-tally( ~Cylinders, data = Cars)
freq_cylinders

# -> 136

# 3. What percentage of cars had 6 cylinders? (Round response to two decimal places)
proportion_cylinders<-tally( ~Cylinders, format="percent", data = Cars)
proportion_cylinders

# -> 44.39

```

4. What proportion of cars had all wheel drive? (Round response to two decimal places)
```{r}
freq_all_wheel_drive<-tally( ~All_wheel_drive, format="percent", data = Cars)
freq_all_wheel_drive

# 21.49
```


### Q 5
How many sports cars were in the sample?

```{r}
freq_sports_cars<-tally( ~Sports, data = Cars)
freq_sports_cars
```

### Q 6 Create a bar chart showing the distribution of the proportion of different total car cylinders in the sample. Save the bar chart as an image and upload it to this exercise
This question was imported from an external source. It was a 'File Upload' question, which is not supported in this quiz tool.
```{r}
proportion_car_cylinders <- barplot(freq, main = "Cars", ylab="Percent", ylim=c(0,50))
dev.copy(png, filename='exports/proportion_car_cylinders.png')
dev.off()
```

### Q 7
Create a contingency table showing the column proportions of cylinders by sports car.
```{r}
contingency_car_cylinders <- tally(~Sports + Cylinders, format="percent", data=Cars)
contingency_car_cylinders
c(4.6728972 + 3.2710280 + 0.2336449 + 0.2336449)

```

What proportion of sports cars have six or more cylinders? (Round answer to two decimal places)

```{r}
proportion_6_contingency_car_cylinders <- round(
  tally(~Sports + Cylinders,  data=Cars, format = "proportion"), 4)
proportion_6_contingency_car_cylinders

c(0.0467 + 0.0327 + 0.0023 + 0.0023)

```

### Q 8
What proportion of non-sports cars have six or more cylinders? (Round answer to two decimal places)

```{r}
c(39.7196262 + 17.0560748 + 0.2336449 + 0.4672897)

```

### Q 9
Are sports cars more likely to have 6 or more cylinders than non-sports cars?
	Yes
	No
```{r}

```

### Upload: Q 10 Create a clustered bar chart comparing the proportion of sports and non-sports car cylinder numbers. Save the plot and upload it to this exercise.



```{r}

cars_clusteredbarchart <-
  tally( ~ cut | clarity, format = "proportion", data=Cars)

barcars <- 
  barplot(table, 
    main = "Cars",
    ylab=~Sports,
    ylim=c(0,.8),
    legend=rownames(table),
    beside=TRUE,
    args.legend=c(x = "top",
      horiz=TRUE,
      title="Sports vs Not"),
      xlab="Clarity",
      height = 100)
barcars
```

```{r}
sports_car_table<-tally( ~ proportion_6_contingency_car_cylinders | !proportion_6_contingency_car_cylinders , format = "proportion", data=Cars)
sports_car_table
```

### Q 11
What is the modal cylinder number for non-sports cars?

```{r}

```

### Q 12
What is the modal cylinder number for sports cars?

```{r}

```

### Q 13 Produce a dot plot of car kilowatts. Save the plot and upload to this question.
This question was imported from an external source. It was a 'File Upload' question, which is not supported in this quiz tool.
```{r}

```

### Q 14
How would you describe the shape of the kilowatts variable depicted in the dot plot?
	Symmetric
	Skewed to the right
	Skewed to the left
	Multi-modal
```{r}

```

### Q 15 Produce a histogram of city fuel economy with 18 bins. Save the plot and upload it to this exercise
This question was imported from an external source. It was a 'File Upload' question, which is not supported in this quiz tool.
```{r}

```

### Q 16
How would you describe the distribution shape of the city fuel economy depicted in the histogram?
	Skewed to the left
	Multi-modal
	Symmetric
	Skewed to the right
```{r}

```

### Q 17
For city fuel economy, would you expect the mean or median to be higher based on the shape of the distribution?
	About the same
	Median would be higher
	Mean would be higher
	Impossible to say
```{r}

```

### Q 18 Produce a histogram of car length with 18 bins. Save the plot and upload it to this exercise.
This question was imported from an external source. It was a 'File Upload' question, which is not supported in this quiz tool.
```{r}

```

### Q 19
How would you describe the distribution shape of car length depicted in the histogram?
	Skewed to the left
	Multi-modal
	Symmetric
	Skewed to the right
```{r}

```

### Q 20 Create a filtered dataset called Cars_filtered, with only 4 and 6 cylinder cars selected. Produce a side-by-side box plot of 4 and 6 cylinder cars on highway fuel economy. Save the plot and upload it to this exercise.
This question was imported from an external source. It was a 'File Upload' question, which is not supported in this quiz tool.
```{r}

```

### Q 21
How many suggested outliers are present for the four cylinder cars?

```{r}

```

### Q 22
Which cylinder, 4 or 6, has the higher IQR for highway fuel economy?
	4
	6
```{r}

```

### Q 23
What is the median highway fuel economy for 4 cylinder cars?

```{r}

```

### Q 24
What is the mean highway fuel economy for 4 cylinder cars? (Round answer to two decimal places)

```{r}

```

### Q 25
What is the standard deviation for highway fuel economy for 4 cylinder cars? (Round answer to two decimal places)

```{r}

```

### Q 26
What is the IQR for highway fuel economy for the 6 cylinder cars?

```{r}

```

### Q 27
What is the upper fence for an outlier on highway fuel economy in the 4 cylinder cars?

```{r}

```

### Q 28
What is the range of highway fuel economy for 6 cylinder cars?

```{r}

```

### Q 29 Produce a scatter plot of city fuel economy by highway fuel economy. Save the plot and upload it to this exercise.
This question was imported from an external source. It was a 'File Upload' question, which is not supported in this quiz tool.
```{r}

```

### Q 30
Which of the following statements best explains the relationship between city and highway fuel economy?
	As highway fuel economy increases, city fuel economy tends to increase.
	As highway fuel economy increases, city fuel economy tends to decrease.
	As there are a few outliers present, it's difficult to determine the nature of the relationship between city and highway fuel economy.
	There is no relationship between city and highway fuel economy
