---
title: "Module 3: Tasks - Models for Stationary Time Series"
---

```{r}
library(TSA)
# set.seed(6452135)
setwd("/Users/phil/code/data-science/cheatsheet/source/includes/time-series/03-models-stationary-time-series/tasks/raw-data")
```

#--- Task 1 --- 

## 1. Simulate 100 observations from the following processes with variance of 1 using R and plot their sample ACF and PACF, and study the characteristics of the sample ACF and PACF to observe the true values of model parameters.


```{r}
n = 100
```

### a) AR(1) with  LaTeX: \phi = 0.6 
ϕ = 0.6
```{r}
sim.data1 = arima.sim(list(order = c(1,0,0), ar = 0.6), n = n) 
acf(sim.data1)
pacf(sim.data1)
# Set of possible models : {AR(1), ARMA(1,2)}
```

### b) AR(1) with  LaTeX: \phi = -0.6 
ϕ = − 0.6
```{r}
sim.data2 = arima.sim(list(order = c(1,0,0), ar = -0.6), n = n) 
acf(sim.data2)
pacf(sim.data2)
# Set of possible models : {ARMA(1,1), AR(1) = ARMA(1,0), MA(1) = ARMA(0,1)} 
```

<!-- ~old_todo~ c) AR(2) with LaTeX: \phi=[0.6, -0.6] ϕ = [ 0.6 , − 0.6 ] -->

### d) MA(1) with LaTeX: \theta = 0.5 
θ = 0.5
```{r}
sim.data3 = arima.sim(list(order = c(0,0,1), ma = 0.95), n = n) 
acf(sim.data3)
pacf(sim.data3)

# Set of possible models : {MA(1), ARMA(2,1), ARMA(3,1), ARMA(4,1)}
```

### e) MA(1) with LaTeX: \theta = -0.5 
θ = − 0.5
```{r}
sim.data4 = arima.sim(list(order = c(0,0,1), ma = -0.5), n = n) 
acf(sim.data4)
pacf(sim.data4)
# Set of possible models : {MA(1), ARMA(1,1)}
```

<!-- ~old_todo~ f) MA(2) with LaTeX: \theta = [-0.5,0.5] θ = [ − 0.5 , 0.5 ] -->

### g) ARMA(1,1) with LaTeX: \phi = 0.6; \theta = 0.5 
ϕ = 0.6 
θ = 0.5
```{r}
sim.data5 = arima.sim(list(order = c(1,0,1), ar = 0.6, ma = 0.5), n = n) 
acf(sim.data5)
pacf(sim.data5)
# Set of possible models : {AR(2), ARMA(2,2), ARMA(2,3), , ARMA(2,4)}
```

### h) ARMA(1,1) with LaTeX: \phi=0.6; \theta=-0.5 
ϕ = 0.6 ; 
θ = − 0.5
<!-- ~old_todo~ is 0.15 a typo? -->
```{r}
sim.data6 = arima.sim(list(order = c(1,0,1), ar = 0.6, ma = -0.15), n = n)
acf(sim.data6)
pacf(sim.data6)
# Set of possible models : {ARMA(1,2), ARMA(1,1)}
```

### i) ARMA(1,1) with LaTeX: \phi=-0.6; \theta=0.5 
ϕ = − 0.6; 
θ = 0.5
```{r}
sim.data7 = arima.sim(list(order = c(1,0,1), ar = -0.6, ma = 0.5), n = n)
plot(sim.data7)
acf(sim.data7)
pacf(sim.data7)
# Set of possible models : {ARMA(1,2), ARMA(1,2)}
```

### j) ARMA(1,1) with LaTeX: \phi=-0.6; \theta=-0.5 
ϕ = − 0.6; 
θ = − 0.5
```{r}
sim.data8 = arima.sim(list(order = c(1,0,1), ar = -0.6, ma = -0.5), n = n)
acf(sim.data8)
pacf(sim.data8)
# Set of possible models : {ARMA(2,2), ARMA(2,3), ARMA(1,2)}
```

<!-- ~old_todo~ ### k) ARMA(2,2) with  LaTeX: \phi=[0.6,-0.6]; \theta=[-0.5,0.5] -->

## 2. Stationary ARMA Models
- Load the simulated datasets into R. 
- Then use time series, sample ACF and PACF plots to infer about stationarity and orders of a possible ARMA model for each dataset.

### a) data.sim1 
```{r fig.height=4, fig.width=6}
#--- True Model: ARMA(2,3)
arma.data1 = read.table(file="./raw-data/data.sim1.csv")
arma.data1 = ts(arma.data1)
plot(arma.data1, type = "o")
acf(arma.data1)
pacf(arma.data1)
# Set of possible models : {ARMA(3,3), ARMA(2,3), ARMA(3,2)}
```

### b) data.sim2 
```{r fig.height=4, fig.width=6}
#--- True Model: AR(3)
arma.data2 = read.table(file="./raw-data/data.sim2.csv")
arma.data2 = ts(arma.data2)
plot(arma.data2, type = "o")
acf(arma.data2)
pacf(arma.data2)
# Set of possible models : {ARMA(2,0), ARMA(3,0)}
```

### c) data.sim3 
```{r fig.height=4, fig.width=6}
#--- True Model: AR(1)
arma.data3 = read.table(file="./raw-data/data.sim3.csv")
arma.data3 = ts(arma.data3)
plot(arma.data3, type = "o")
acf(arma.data3)
pacf(arma.data3)
# Set of possible models : {AR(1)}
```

### d) data.sim4 
```{r fig.height=4, fig.width=6}
#--- True Model: ARMA(2,4)
arma.data4 = read.table(file="./raw-data/data.sim4.csv")
arma.data4 = ts(arma.data4)
plot(arma.data4, type = "o")
acf(arma.data4)
pacf(arma.data4)
# Set of possible models : {ARMA(2,4) ARMA(2,2)}
```

### e) data.sim5 
```{r fig.height=4, fig.width=6}
#--- True Model: MA(2)
arma.data5 = read.table(file="./raw-data/data.sim5.csv")
arma.data5 = ts(arma.data5)
plot(arma.data5, type = "o")
acf(arma.data5)
pacf(arma.data5)
# Set of possible models : {ARMA(2,2), ARMA(1,2), ARMA(2,1), ARMA(1,1)}
```
