---
title: "MATH2349 Semester 1, 2018"
author: "Phil Steinke s3725547@student.rmit.edu.au"
subtitle: "Assignment 1 - Victorian family violence cases 2012-2017"
output:
  html_notebook: default
---

## Setup
```{r, echo = FALSE, warnings = FALSE}

library(readr) # Useful for importing data
library(foreign) # Useful for importing SPSS, SAS, STATA etc. data files
library(rvest) # Useful for scraping HTML data
library(knitr) # Useful for creating nice tables
library(forcats)
library(mosaic)
library(readxl)
library(rmarkdown)
library("miceadds", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("plotly")



```

## Data Description

"Victims Support Agency Data Tables- 2016-17.xlsx"
Table 2. Number of VAP family violence cases initiated for new clients by client gender and age group, July 2012 to June 2017

Source: [https://www.crimestatistics.vic.gov.au/sites/default/files/embridge_cache/emshare/original/public/2017/12/74/906ab3fb8/Victims%20Support%20Agency%20Data%20Tables-%202016-17.xlsx](https://www.crimestatistics.vic.gov.au/sites/default/files/embridge_cache/emshare/original/public/2017/12/74/906ab3fb8/Victims%20Support%20Agency%20Data%20Tables-%202016-17.xlsx)

```{r}
#' VICTIMS SUPPORT AGENCY DATA TABLES- 2016-17.XLSX
#' TABLE 2. NUMBER OF VAP FAMILY VIOLENCE CASES INITIATED FOR NEW CLIENTS BY CLIENT GENDER AND AGE GROUP, 
#' JULY 2012 TO JUNE 2017
#'
#' VAP FAMILY VIOLENCE CASES INITIATED FOR NEW CLIENTS BY CLIENT GENDER AND AGE GROUP
#' 
#' @format Starting format xlxs with 53 observations/rows and 7 variables/cols
#' only 46 observations/rows and 7 variables/cols are imported to not include whitespace and the totals
#' \describe{
#'   \item{\code{Gender}}{character. GENDER OF CLIENT THAT REPORTED FAMILY VIOLENCE. LEVELS: MALE/FEMALE.}
#'   \item{\code{age group}}{character. DESCRIPTION.}
#'   \item{\code{2012-13}}{character. COUNT OF REPORTED INCIDENTS FROM JULY 2012 TILL JUNE 2013.}
#'   \item{\code{2013-14}}{character. COUNT OF REPORTED INCIDENTS FROM JULY 2013 TILL JUNE 2014.}
#'   \item{\code{2014-15}}{character. COUNT OF REPORTED INCIDENTS FROM JULY 2014 TILL JUNE 2015.}
#'   \item{\code{2015-16}}{character. COUNT OF REPORTED INCIDENTS FROM JULY 2015 TILL JUNE 2016.}
#'   \item{\code{2016-17}}{character. COUNT OF REPORTED INCIDENTS FROM JULY 2016 TILL JUNE 2017.}
#' }
#' "Victims Support Agency Data Tables- 2016-17.xlsx"
#'
#' @format end format is a dataframe with 28 observations/rows and 7 variables/cols
#' \describe{
#'   \item{\code{Gender}}{character. GENDER OF CLIENT THAT REPORTED FAMILY VIOLENCE. LEVELS: MALE/FEMALE. }
#'   \item{\code{Age Range}}{character. AGE OF PARTICIPANTS DIVIDED INTO 5 YEAR INCRIMENTS.}
#'   \item{\code{2012-13}}{integer. COUNT OF REPORTED INCIDENTS FROM JULY 2012 TILL JUNE 2013.}
#'   \item{\code{2013-14}}{integer. COUNT OF REPORTED INCIDENTS FROM JULY 2013 TILL JUNE 2014.}
#'   \item{\code{2014-15}}{integer. COUNT OF REPORTED INCIDENTS FROM JULY 2014 TILL JUNE 2015.}
#'   \item{\code{2015-16}}{integer. COUNT OF REPORTED INCIDENTS FROM JULY 2015 TILL JUNE 2016.}
#'   \item{\code{2016-17}}{integer. COUNT OF REPORTED INCIDENTS FROM JULY 2016 TILL JUNE 2017.}
#' }
"family_violence"
```

As a minimum, your data set should include:
* one numeric variable = number of family violence cases per year
* one qualitative (categorical) variable = Age Range

This dataset show most reported assaults with women occur between the ages of 25-49. Each age bracket within that range (of 5 years) have approximately double the reported assaults of children and teenagers. The data also shows an increase in reported assaults per year over the last 5 years. 

## Read/Import Data

```{r}
rm(list=ls())
setwd("~/code/tldr/data-science/data-preprocessing-math2349/assignment1/data/")

# Read/Import the data into R, then save it as a data frame.

family_violence <-
  read_excel(
    "Victims Support Agency Data Tables- 2016-17.xlsx",
    sheet = "Table 2",
    range = cell_rows(12:58)
  ) %>%
  data.frame()

# `stringsAsFactors = FALSE` wont work here, so I set it in my defaults

class(family_violence) # -> family violence is a "data.frame"

# You must also provide the R codes with outputs
head(family_violence)

```

## Inspect and Understand

### * check the dimensions of the data frame.
```{r}
dim(family_violence)
# OR
nrow(family_violence)
ncol(family_violence)

# check the attributes in the data.
attributes(family_violence)

```
- It has 46 rows and 7 columns
- It's names are X__1, ...
- It's row names are numbers 1,2,3...
- It's a data.frame

## check the data types (i.e., character, numeric, integer, factor, and logical) of the variables in the data set.

```{r}

class(family_violence) # -> family_violence is a data.frame

class(family_violence[,1]) # -> "Gender and age group" is a character

head(family_violence[,2]) # [1] NA             "0 - 4"        "5 - 9"        "10 - 14"      "15 - 19"      "20 - 24"
class(family_violence[,2]) # -> is a character

head(family_violence[,3]) #  [1] "2012-13" "74"      "84"      "72"      "52"      "64"      "62"         
class(family_violence[,3])# ->  is a character

head(family_violence[,4]) #  [1] "2013-14" "61"      "121"     "80"      "70"      "47"      "60"          
class(family_violence[,4])  # ->  is a character

head(family_violence[,5]) #  [1] "2013-14" "61"      "121"     "80"      "70"      "47"      "60"          
class(family_violence[,5])  # ->  is a character

head(family_violence[,6]) # "2015-16" "41"      "107"     "88"      "74"      "95"     
class(family_violence[,6])  # ->  is a character

```

- Everything is treated as a character because of the column titles are included in the spreadsheet

### check the levels of factor variables

```{r}
# family_violence[1,] # column names for reference

levels_gender <-
  c(family_violence[,1]) %>%
  factor(ordered= TRUE) %>%
  levels() %>%
  print()

levels_age_range <-
  c(family_violence[,2]) %>%
  factor(ordered= TRUE) %>%
  levels() %>%
  print()

cat("\nLevels for all year cols from 2012-17\n including titles")
levels_all_years <-
c(family_violence[,3],
  family_violence[,4],
  family_violence[,5],
  family_violence[,6],
  family_violence[,7]
) %>%
  factor() %>%
  levels() %>%
  print()

cat("\nLevels from 2012-13\n")
levels_2012_13 <-
  c(family_violence[,3]) %>%
  factor(ordered= TRUE) %>%
  levels() %>%
  print()

```

# * check the column names in the data frame, rename them if required.

```{r}
# check the column names in the data frame
colnames(family_violence)

```

```{r}
# rename them if required.
colnames(family_violence) <- c("Gender", "Age Range", c(family_violence[1,3:7]))

#The excel doesn't include Male/Female accross all of the fields, so here I've filled them in:
family_violence[c(3:16),1] <- "Male"
family_violence[c(18:31),1] <- "Female"

# Removing the empty rows and rows with totals in them
family_violence <- family_violence[-c(1, 16, 31:46), ]

# Fixing the Row numbering
rownames(family_violence) <- c(1:length(family_violence$`Gender`))
family_violence


```

```{r}
class(family_violence) # -> family_violence is a data.frame
family_violence[1, 'Age Range'] # -> "0 - 4"
class(family_violence[3, 'Age Range']) # -> "Age Range" is a character
family_violence[3, '2012-13']
class(family_violence[3, '2012-13']) # -> "Year" is a character
family_violence[1, 4]
class(family_violence[1, 4]) # -> "Gender" and N/A is a character

```


```{r}
# fixing the data types: rename/rearrange if required

cat("Setting each year's data to integers\n")
class(family_violence[3:7])
family_violence[3:7] <- Map(as.integer, family_violence[3:7])

Map(is.integer, family_violence[3:7])

# Previous code that seemed cumbersome:
#class(family_violence$`2012-13`)
family_violence$`2012-13` %>%
  as.integer() -> family_violence$`2012-13`
#class(family_violence$`2012-13`)

cat("\nLevels for all years again: from 2012-17\n including titles")
levels_all_years <-
c(family_violence[,3],
  family_violence[,4],
  family_violence[,5],
  family_violence[,6],
  family_violence[,7]
) %>%
  factor() %>%
  levels() %>%
  print()

```

## New data types tests
```{r}
cat("New data types\n")
class(family_violence) # -> family_violence is a data.frame
cat("Age Range\n")
family_violence[1, 'Age Range'] # -> "0 - 4"
class(family_violence[3, 'Age Range']) # -> "Age Range" is a character
cat("Year col 2012-13\n")
family_violence$'2012-13'
class(family_violence$'2012-13') # -> All Year cols are now an integer
cat("single value from a year column 2012-13\n")
family_violence[1, 5]
class(family_violence[1, 4]) # -> Grabbing a single value from a year col which is now an integer
dim(family_violence)

```

## Subsetting I

Subset the data frame using first 10 observations (include all variables). Then convert it to a matrix.

```{r}
# Subset the data frame using first 10 observations (include all variables)

# What are all variables?
names(family_violence) -> all_variables
all_variables

# I assume you mean this because all_variables
data_frame_subset <- family_violence[1:10,]
data_frame_subset

# Then convert it to a matrix

data_frame_subset %>%
  as.matrix(
  ) %>%
  print()

data_frame_matrix1 <- data.matrix(data_frame_subset, rownames.force = NA)
class(data_frame_matrix1) # Matrix
data_frame_matrix2 <- as.matrix(data_frame_subset)
class(data_frame_matrix2) # Matrix
data_frame_matrix3 <- apply(data_frame_subset, 2, as.matrix) 
class(data_frame_matrix3) # Matrix Trinity

```

## Subsetting II

```{r}
## Subset the data frame including only first and the last variable in the data set

# Grabbing the variables:
names(family_violence) -> all_variables
all_variables
```

```{r}

family_violence %>%
  subset (
    select = c(
      1,
      length(family_violence)
    )
  ) -> first_and_last_subset
head(first_and_last_subset)

```

## save it as an R object file (.RData).
#### This didn't work:
```{r}

save.image() # Saving the workspace

first_and_last_subset

save(first_and_last_subset, file = "data/first_and_last_subset.Rdata")
rm(first_and_last_subset)
testing_save_worked <- load("data/first_and_last_subset.Rdata")

identical(first_and_last_subset, testing_save_worked) # FALSE

```

```{r}
# Using load.Rdata2 from miceadds instead:

save.Rdata(first_and_last_subset, "data/first_and_last_subset.RData")

testing_save_worked <- load.Rdata2(filename = "data/first_and_last_subset.RData", path=getwd())

identical(first_and_last_subset, testing_save_worked) # [1] TRUE
```
<br>
<br>
